---
title: "Palmer Drought Severity Index"
author: "AnnaClaire Marley"
date: "7/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
source("Functions/util.R")
source('Functions/graphing.R')
```

```{r, include = FALSE}
# read in NN whole PDSI

nn_pdsi <- read_csv("NN_whole/nn_pdsi_79-19.csv") %>% 
  rename(date = DateTime,
         pdsi = `Palmer Drought Severity Index (PDSI) (gridMET) at 4005, 1979-01-01 to 2019-06-09`) %>% 
  add_water_year()


```


```{r, include = FALSE, warning=FALSE}

# read in the agency data & add water year
chinle_pdsi <- read_csv("Agencies/chinle_pdsi.csv") 
east_navajo_pdsi <- read_csv("Agencies/east_navajo_pdsi.csv")
west_navajo_pdsi <- read_csv("Agencies/west_navajo_pdsi.csv")
shiprock_pdsi <- read_csv("Agencies/shiprock_pdsi.csv")
fort_defiance_pdsi <- read_csv("Agencies/fort_defiance_pdsi.csv")

# winter months
chinle_pdsi_wint <- chinle_pdsi %>% 
  filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

east_navajo_pdsi_wint <- east_navajo_pdsi %>% 
  filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

west_navajo_pdsi_wint <- west_navajo_pdsi %>% 
  filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

shiprock_pdsi_wint <- shiprock_pdsi %>% 
  filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

fort_defiance_pdsi_wint <- fort_defiance_pdsi %>% 
  filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

```

```{r, include = FALSE, warning=FALSE}
# take the average yearly pdsi value 
chinle_an_av <- calc_pdsi_metrics(chinle_pdsi, "year", mean)
east_navajo_an_av <- calc_pdsi_metrics(east_navajo_pdsi, "year", mean)
west_navajo_an_av <- calc_pdsi_metrics(west_navajo_pdsi, "year", mean)
shiprock_an_av <- calc_pdsi_metrics(shiprock_pdsi, "year", mean)
fort_defiance_an_av <- calc_pdsi_metrics(fort_defiance_pdsi, "year", mean)

# take the monthly mean of the yearly data

chinle_mnth_av <- calc_pdsi_metrics(chinle_pdsi, "month", mean)
east_navajo_mnth_av <- calc_pdsi_metrics(east_navajo_pdsi, "month", mean)
west_navajo_mnth_av <- calc_pdsi_metrics(west_navajo_pdsi, "month", mean)
shiprock_mnth_av <- calc_pdsi_metrics(shiprock_pdsi, "month", mean)
fort_defiance_mnth_av <- calc_pdsi_metrics(fort_defiance_pdsi, "month", mean)

# take the monthly mean of the winter data
chinle_mnth_av_wint <- calc_pdsi_metrics(chinle_pdsi_wint, "month", mean)
east_navajo_mnth_av_wint <- calc_pdsi_metrics(east_navajo_pdsi_wint, "month", mean)
west_navajo_mnth_av_wint <- calc_pdsi_metrics(west_navajo_pdsi_wint, "month", mean)
shiprock_mnth_av_wint <- calc_pdsi_metrics(shiprock_pdsi_wint, "month", mean)
fort_defiance_mnth_av_wint <- calc_pdsi_metrics(fort_defiance_pdsi_wint, "month", mean)

# visualize example 
ggplot(chinle_mnth_av_wint, aes(x = date, y = pdsi)) +
  geom_col() +
  theme_classic()

```


#### Correlation of PDSI monthly averages for the Agencies
```{r, echo = FALSE, warning=FALSE, message=FALSE}
## look to see how correlated the PDSI is between regions

pdsi_agencies <- chinle_mnth_av %>% 
  rename(chinle = pdsi) %>% 
  left_join(east_navajo_mnth_av, by = "date") %>% 
  rename(east_navajo = pdsi) %>% 
  left_join(west_navajo_mnth_av, by = "date") %>% 
  rename(west_navajo = pdsi) %>% 
  left_join(shiprock_mnth_av, by = "date") %>% 
  rename(shiprock = pdsi) %>% 
  left_join(fort_defiance_mnth_av, by = "date") %>% 
  rename(fort_defiance = pdsi) %>% 
  select(date, chinle, east_navajo, west_navajo, shiprock, fort_defiance) 
  
pdsi_cor_df<- pdsi_agencies %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "date")

# correlation matrix
cor_pdsi <- cor(pdsi_cor_df)
corrplot(cor_pdsi, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)

```

- All the agencies have a high PDSI correlation


```{r, echo=FALSE, warning=FALSE, message = FALSE}
#### Now look to see if there are correlations in PDSI and swe levels ####

# first read in Chuska swe data and calculate metrics

chuska_swe <- read_csv("winter_wy/ch_swe_wy_03_19.csv")
chuska_swe_mnth_av <- calc_swe_metrics(chuska_swe, "month", mean)
chuska_swe_an_av <- calc_swe_metrics(chuska_swe, "year", mean)
chuska_mnth_anom <- calc_month_anom(chuska_swe_mnth_av) %>% 
  arrange(date)

# plot swe vs pdsi

ship_ch_plot <- plot(shiprock_mnth_av_wint$pdsi, chuska_mnth_anom$anomaly)
ship_ch_plot_cor <- cor.test(shiprock_an_av$pdsi, chuska_swe_an_av$swe_mm)
```


### Correlation between Chuska monthly average swe and the monthly average PDSI of the agencies
```{r, echo=FALSE, warning=FALSE, message = FALSE}
## see how chuska correlates with the pdsi of the agencies

agencies_v_chuska <- pdsi_agencies %>% 
  left_join(chuska_swe_mnth_av, by = "date") %>% 
  select(-waterYear) %>% 
  filter(swe_mm != "NA") %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "date")

cor_pdsi_chuska <- cor(agencies_v_chuska)
corrplot(cor_pdsi_chuska, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)
```

- not a very high correlation

```{r, echo = FALSE, warning=FALSE, message = FALSE}

## make dataframes with swe and pdsi for shiprock

chuska_ship <- shiprock_mnth_av %>% 
  left_join(chuska_mnth_anom, by = "date") %>% 
  select(date, pdsi, anomaly, sign.x, waterYear.x, anomaly_perc, sign.y) 

## make graph of the two
chuska_ship_graph <- chuska_ship %>% 
  ggplot(aes(x = date)) +
  geom_col(aes(y = anomaly_perc/100, fill = sign.y), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  labs(
    y = "SWE Anomaly (scaled down by 10)"
  ) +
  geom_col(aes(y = pdsi), color = "#1219cc", alpha = 0.5) +
  scale_y_continuous(sec.axis = sec_axis(~., name = "PDSI"), expand = c(0,0)) +
  theme_classic()

chuska_ship_graph



```

```{r, echo = FALSE, message = FALSE}
# chuska vs chinle

chuska_chin <- chinle_mnth_av %>% 
  left_join(chuska_mnth_anom, by = "date") %>% 
  select(date, pdsi, anomaly, sign.x, waterYear.x, sign.y) 

## make graph of the two
chuska_chin_graph <- chuska_chin %>% 
  ggplot(aes(x = date)) +
  geom_col(aes(y = anomaly / 10, fill = sign.y), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  labs(
    y = "SWE Anomaly (scaled down by 10)"
  ) +
  geom_col(aes(y = pdsi), color = "#1219cc", alpha = 0.5) +
  scale_y_continuous(sec.axis = sec_axis(~., name = "PDSI"), expand = c(0,0)) +
  theme_classic()

chuska_chin_graph

```

```{r, echo = FALSE}
#fortdefiance anomaly

fort_def_mnth_intra_av <- fort_defiance_mnth_av %>% 
    mutate(month = month(date)) %>% 
    group_by(month) %>% 
    summarize(mean_pdsi = mean(pdsi))
    
fort_def_anom  <- fort_defiance_mnth_av %>% 
    mutate(month = month(date)) %>% 
    merge(fort_def_mnth_intra_av, by = "month")%>%
    mutate(anomaly_pdsi = pdsi - mean_pdsi) %>% 
    mutate(pdsi_anom_perc = (anomaly_pdsi/mean_pdsi)*100)

# chuska vs fort defiance
chuska_fd <- fort_def_anom %>% 
  left_join(chuska_mnth_anom, by = "date") %>% 
  select(date, pdsi, anomaly, sign.x, waterYear.x, anomaly_pdsi, pdsi_anom_perc, anomaly_perc, sign.y) 

## make graph of the two
chuska_fd_graph <- chuska_fd %>% 
  ggplot(aes(x = date)) +
  geom_col(aes(y = anomaly_perc, fill = sign.y), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  labs(
    y = "SWE % Difference from Average"
  ) +
  geom_line(aes(y = anomaly_pdsi*100), color = "#1219cc", alpha = 0.5) +
  scale_y_continuous(sec.axis = sec_axis(~., name = "PDSI % Difference from Average"), expand = c(0,0)) +
  theme_classic()

chuska_fd_graph

chuska_fd_graph2 <- chuska_fd %>% 
  ggplot(aes(x = date)) +
  geom_col(aes(y = anomaly_perc, fill = sign.y), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  labs(
    y = "SWE % Difference from Average",
    x = "Water Year"
  ) +
  geom_line(aes(y = anomaly_pdsi*100), color = "#1219cc", alpha = 0.5) +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "PDSI Anomaly (scaled x 100)"), expand = c(0,0)) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_classic() +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
        axis.text.y.right = element_text(color = "#1219cc"),
        axis.title.y.right = element_text(color = "#1219cc"))

#chuska_fd_graph2


```

```{r, echo = FALSE, error = TRUE}
# Chuska and fort defiance agency % difference from average correlations

plot(chuska_fd$anomaly_perc, chuska_fd$anomaly_pdsi)

cor.test(chuska_fd$anomaly_perc, chuska_fd$anomaly_pdsi)
```


```{r, echo = FALSE, warning=FALSE}
## quick make chuska anomaly graph for poster

 # first make variables for ease of graphing  
      startWaterYear = min(chuska_mnth_anom$waterYear)
      endWaterYear = max(chuska_mnth_anom$waterYear)
      # Create sequence of real dates, on which to place the water year lables
      breakDates = seq.Date(from = ymd(paste(startWaterYear,1,1, sep='-')), length.out = endWaterYear - startWaterYear + 1, by = "1 year")

ch_poster <- ggplot(chuska_mnth_anom, aes(x = date, y = anomaly)) +
  geom_col(aes(y = anomaly / 10, fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_date(breaks= breakDates, labels=seq(startWaterYear,endWaterYear,by=1)) +
  labs(
    y = "SWE Anomaly (mm)",
    x = "Water Year"
  ) +
  theme_classic()

```


### PDSI Whole Navajo Nation
```{r, echo = FALSE, warning=FALSE}
# do some analysis of the whole NN

nn_wk_pdsi <- calc_pdsi_metrics(nn_pdsi, "1 week", mean)

nn_wk_pdsi_graph <- nn_wk_pdsi %>% 
  ggplot(aes(x = date, y = pdsi)) + 
  geom_line() +
  theme_classic()


```

```{r, echo = FALSE, warning=FALSE}
nn_mnth_pdsi <- calc_pdsi_metrics(nn_pdsi, "month", mean)

plot_pdsi(nn_mnth_pdsi, title = "Navajo Nation")

nn_mnth_pdsi <- nn_mnth_pdsi %>% 
  filter(pdsi != "NA") 
  
nn_mnth_pdsi_lm <- lm(nn_mnth_pdsi$pdsi ~ nn_mnth_pdsi$date)
summary(nn_mnth_pdsi_lm)
```

#### Compare PDSI whole NN to total swe for all high elevation regions

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# read in the total weekly swe for the mountain ranges 
mtns_swe <- read_csv("NN_whole/nn_mtns_swe_total.csv")

# calculate monthly anomaly
mtns_swe_anom <- mtns_swe %>% 
  rename(swe_mm = sum_mts_swe_m3) %>% 
  add_water_year() %>% 
  calc_swe_metrics("month", mean) %>% 
  calc_month_anom() 
```

```{r, echo = FALSE, warning=FALSE}
#### graph percent difference from average swe against pdsi

#first merge both things into the same dataframe
mtns_swe_nn_pdsi <- left_join(nn_mnth_pdsi, mtns_swe_anom, by = "date") %>% 
  filter(date >= "2003-10-01") %>% 
  select(date, waterYear.x, anomaly_perc, sign.x,  pdsi, sign.y) %>% 
  rename(pdsi_sign = sign.x,
         swe_sign = sign.y)

# now graph it

nn_pdsi_swe <- mtns_swe_nn_pdsi %>% 
  ggplot(aes(x = date)) + 
  geom_col(aes(y = pdsi, fill = pdsi_sign), show.legend = FALSE) +
  labs(
    x = "Year",
    y = "PDSI",
    title = "Navajo Nation Palmer Drought Severity Index (PDSI) "
  ) +
  scale_x_date(expand = c(0,0)) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  geom_col(aes(y = anomaly_perc/100, alpha = 0.5), fill = "#1219cc") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "SWE Anomaly (scaled down by 100)"), expand = c(0,0)) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_classic() +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
        axis.text.y.right = element_text(color = "#1219cc"),
        axis.title.y.right = element_text(color = "#1219cc")) 
  

nn_pdsi_swe

```


```{r, echo = FALSE, warning=FALSE}

nn_pdsi_swe2 <- mtns_swe_nn_pdsi %>% 
  ggplot(aes(x = date)) + 
  geom_col(aes(y = anomaly_perc/100), fill = "#9dc6e0", show.legend = FALSE) +
  labs(
    x = "Year",
    y = "SWE Percent different from average (scaled down by 100)",
    title = "Navajo Nation Palmer Drought Severity Index (PDSI) "
  ) +
  scale_x_date(expand = c(0,0)) +
  geom_line(aes(y = pdsi)) +
  scale_y_continuous(sec.axis = sec_axis(~., name = "PDSI"), expand = c(0,0)) +
  #scale_color_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_classic() +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
        axis.text.y.right = element_text(color = "#1219cc"),
        axis.title.y.right = element_text(color = "#1219cc")) 
  

nn_pdsi_swe2

```

```{r, echo = FALSE, warning=FALSE}
# mountain ranges winter swe and all of navajo nation AMJ pdsi
mtns_nn_wint_amj <- compare_swe_pdsi(mtns_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                     nn_mnth_pdsi, pdsi_mnths = c(4,5,6))
plot_cor_swe_pdsi(mtns_nn_wint_amj, title = "Winter SWE and NN AMJ PDSI")

mtns_nn_wint_amj_plot <- mtns_nn_wint_amj %>% 
  mutate(mean_swe = mean_swe/100) %>% 
  select(waterYear.x, mean_swe, mean_pdsi) %>% 
  melt(id.vars = "waterYear.x") %>% 
  rename(metric = variable) %>% 
  ggplot(aes(x = waterYear.x, y = value, fill = metric)) +
  geom_col(position = "dodge") +
  theme_classic()

```


### Now let's looks at correlations between different months

#### Notation:
- Winter SWE: November - February (w)
- Spring SWE: March - April (sp)
- Early summer PDSI: May - June (esu)
- Mid summer PDSI: June - July (msu)
- Late summer PDSI: July - August (lsu)


```{r, echo = FALSE, warning=FALSE}

# first clean up the monthly anomaly of the mts a bit the mountains
mtns_swe_anom_clean <- mtns_swe_anom %>% 
  select(date, anomaly_perc, waterYear)

# get the monthly mean of pdsi & filter to 2004
nn_pdsi_mnth <- calc_pdsi_metrics(nn_pdsi, "month", mean) %>% 
  filter(waterYear >= 2004)

```

```{r, echo = FALSE, warning=FALSE}
#### dataframes of combined ### 

# spring swe, early summer pdsi:
swe_pdsi_sp_esu <- compare_swe_pdsi(mtns_swe_anom_clean, swe_mnths = c(3,4), nn_pdsi_mnth, pdsi_mnths = c(5,6))

# spring swe, mid summer pdsi:
swe_pdsi_sp_msu <- compare_swe_pdsi(mtns_swe_anom_clean, swe_mnths = c(3,4), nn_pdsi_mnth, pdsi_mnths = c(6,7))

# spring swe, late summer pdsi:
swe_pdsi_sp_lsu <- compare_swe_pdsi(mtns_swe_anom_clean, swe_mnths = c(3,4), nn_pdsi_mnth, pdsi_mnths = c(7,8))
```


```{r, echo = FALSE, warning=FALSE}
# correlations
plot_cor_swe_pdsi(swe_pdsi_sp_esu, title = "Spring SWE and early Summer PDSI")
plot_cor_swe_pdsi(swe_pdsi_sp_msu, title = "Spring SWE and mid Summer PDSI")
plot_cor_swe_pdsi(swe_pdsi_sp_lsu, title = "Spring SWE and late Summer PDSI")


```

### Compare SWE of Chuska to PDSI watersheds

```{r, echo = FALSE, warning=FALSE}
# read in chuska and extract anomalies

chuska_swe <- read_csv("Chuska/ch_swe_wint_wy_03_19.csv")

# chuska swe monthly anomaly
ch_swe_anom <- chuska_swe %>% 
  calc_swe_metrics("month", mean) %>% 
  calc_month_anom() %>% 
  select(date, waterYear, anomaly_perc, sign) %>% 
  arrange(date)

# Chuska average winter swe each year
ch_swe_av_wint <- chuska_swe %>% 
  group_by(waterYear) %>% 
  summarize(mean_wint_swe = mean(swe_mm))

winter_av <- mean(ch_swe_av_wint$mean_wint_swe)

# Chuska winter anomaly each water Year
ch_swe_anom_wint <- ch_swe_av_wint %>% 
  mutate(wint_anom = mean_wint_swe - winter_av) %>% 
  mutate(anomaly_perc = wint_anom/winter_av)

# now let's find what the standard deviation is for winter swe so that we can 
# use it to define a "dry" vs "wet year"

ch_wint_sd <- sd(ch_swe_anom_wint$anomaly_perc)

# now add a column saying whether or not the anomaly_perc is outwidth one standard deviation
ch_swe_anom_wint <- ch_swe_anom_wint %>% 
  mutate(whether_sd = ifelse(abs(anomaly_perc) >= ch_wint_sd, "yes", "no"))
```


```{r, echo = FALSE, warning=FALSE}
### read in all the watershed data###

# Chaco

chaco_pdsi <- read_csv("watersheds/chaco_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

# Chinle

chinle_pdsi <- read_csv("watersheds/chinle_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

# Upper Puerco
up_puerco_pdsi <- read_csv("watersheds/upper_puerco_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

# middle san juan
mid_sj_pdsi <- read_csv("watersheds/middle_san_juan_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()


```


#### Watershed PDSI timeseries (1981-present)
```{r, echo = FALSE, warning=FALSE}
# visualize PDSI time series

chaco_pdsi_mnth <- calc_pdsi_metrics(chaco_pdsi, "month", mean) 
  plot_pdsi(chaco_pdsi_mnth, title = "Chaco")

chinle_pdsi_mnth <- calc_pdsi_metrics(chinle_pdsi, "month", mean) 
  plot_pdsi(chinle_pdsi_mnth, title = "Chinle")

up_puerco_pdsi_mnth <- calc_pdsi_metrics(up_puerco_pdsi, "month", mean) 
  plot_pdsi(up_puerco_pdsi_mnth, title = "Upper Puerco")

mid_sj_pdsi_mnth <- calc_pdsi_metrics(mid_sj_pdsi, "month", mean)
  plot_pdsi(mid_sj_pdsi_mnth, title = "Middle San Juan")

```

```{r, echo = FALSE, warning=FALSE}

# combine watersheds PDSI and Chuska swe into one dataframe
# shows monthly mean pdsi and monthly mean swe anomaly

ch_chaco <- combine_swe_pdsi(chaco_pdsi, ch_swe_anom)

ch_chinle <- combine_swe_pdsi(chinle_pdsi, ch_swe_anom)

ch_up_puerco <- combine_swe_pdsi(up_puerco_pdsi, ch_swe_anom)

ch_mid_sj <- combine_swe_pdsi(mid_sj_pdsi, ch_swe_anom)

```


#### Chuska SWE Anomalies vs watershed PDSI
```{r, echo = FALSE, warning=FALSE, message = FALSE}
# Visually compare to Chuska SWE

plot_pdsi_swe(ch_chaco, title = "Chaco")

plot_pdsi_swe(ch_chinle, title = "Chinle")

plot_pdsi_swe(ch_up_puerco, title = "Upper Puerco")

plot_pdsi_swe(ch_mid_sj, title = "Middle San Juan")


```

#### Compare Chuska SWE different months of SWE to different months of PDSI
```{r, echo = FALSE, warning=FALSE, message = FALSE}

## first pick out months of interest
ch_chaco_sp_esu <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(3,4), chaco_pdsi_mnth, pdsi_mnths = c(5,6))
ch_chinle_sp_esu <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(3,4), chinle_pdsi_mnth, pdsi_mnths = c(5,6))
ch_up_puerco_sp_esu <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(3,4), up_puerco_pdsi_mnth, pdsi_mnths = c(5,6))
ch_mid_sj_sp_esu <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(3,4), mid_sj_pdsi_mnth, pdsi_mnths = c(5,6))

```

```{r, echo = FALSE, warning=FALSE}
# now graph the relationships
plot_cor_swe_pdsi(ch_chaco_sp_esu, title = "Spring SWE and Chaco early Summer PDSI")
plot_cor_swe_pdsi(ch_chinle_sp_esu, title = "Spring SWE and Chinle early Summer PDSI")
plot_cor_swe_pdsi(ch_up_puerco_sp_esu, title = "Spring SWE and Upper Puerco early Summer PDSI")
plot_cor_swe_pdsi(ch_mid_sj_sp_esu, title = "Spring SWE and Mid San Juan early Summer PDSI")


```

### Winter SWE Chuskas and AMJ PDSI watersheds
```{r, echo = FALSE, warning=FALSE, message = FALSE}

## winter swe vs AMJ PDSI
ch_chaco_wint_amj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                    chaco_pdsi_mnth, pdsi_mnths = c(4,5,6))
ch_chinle_wint_amj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                     chinle_pdsi_mnth, pdsi_mnths = c(4,5,6))
ch_up_puerco_wint_amj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                        up_puerco_pdsi_mnth, pdsi_mnths = c(4,5,6))
ch_mid_sj_wint_amj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4),
                                     mid_sj_pdsi_mnth, pdsi_mnths = c(4,5,6))

```

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# now graph the relationships
plot_cor_swe_pdsi(ch_chaco_wint_amj, title = "Winter SWE and Chaco AMJ PDSI")
plot_cor_swe_pdsi(ch_chinle_wint_amj, title = "Winter SWE and Chinle AMJ PDSI")
plot_cor_swe_pdsi(ch_up_puerco_wint_amj, title = "Winter SWE and Upper Puerco AMJ PDSI")
plot_cor_swe_pdsi(ch_mid_sj_wint_amj, title = "Winter SWE and Mid San Juan AMJ PDSI")


```

- generally pretty low positive correlations between winter swe and AMJ PDSI

Mid San Juan and Chuska have the highest correlation of winter swe and AMJ PDSI. Let's check it out

```{r, echo = FALSE, message = FALSE}

ch_mid_sj_wint_amj_plot <- ch_mid_sj_wint_amj %>% 
  mutate(mean_swe = mean_swe/10) %>% 
  select(waterYear.x, mean_swe, mean_pdsi) %>% 
  melt(id.vars = "waterYear.x") %>% 
  rename(metric = variable) %>% 
  ggplot(aes(x = waterYear.x, y = value, fill = metric)) +
  geom_col(position = "dodge") +
  theme_classic()

ch_mid_sj_wint_amj_plot

```

- 3/8 negative swe anomaly years followed by negative pdsi AMJ
- 5/8 positive swe anomaly years followed by positive pdsi AMJ

### Chuska winter swe and watershed PDSI January - June (6 month PDSI)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
ch_chaco_wint_jj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                    chaco_pdsi_mnth, pdsi_mnths = c(1,2,3,4,5,6))
ch_chinle_wint_jj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                     chinle_pdsi_mnth, pdsi_mnths = c(1,2,3,4,5,6))
ch_up_puerco_wint_jj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                        up_puerco_pdsi_mnth, pdsi_mnths = c(1,2,3,4,5,6))
ch_mid_sj_wint_jj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4),
                                     mid_sj_pdsi_mnth, pdsi_mnths = c(1,2,3,4,5,6))

```

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# now graph the relationships
plot_cor_swe_pdsi(ch_chaco_wint_jj, title = "Winter SWE and Chaco Jan-June PDSI")
plot_cor_swe_pdsi(ch_chinle_wint_jj, title = "Winter SWE and Chinle Jan-JunePDSI")
plot_cor_swe_pdsi(ch_up_puerco_wint_jj, title = "Winter SWE and Upper Puerco Jan-June PDSI")
plot_cor_swe_pdsi(ch_mid_sj_wint_jj, title = "Winter SWE and Mid San Juan Jan-June PDSI")


```

- Generally higher correlations with winter swe and 6 month PDSI then there were for AMJ PDSI
- Upper Puerco has a correlation of almost 0.4

```{r, include = FALSE}
# just need to quickly grab the 6 month pdsi value for each watershed for comparing to spi
chaco_pdsi_jj <- ch_chaco_wint_jj %>% 
  select(waterYear.y, mean_pdsi) %>% 
  rename(waterYear = waterYear.y)

chinle_pdsi_jj <- ch_chinle_wint_jj %>% 
  select(waterYear.y, mean_pdsi) %>% 
  rename(waterYear = waterYear.y)


```


LEt's look at upper puerco and chuska on a timeline for Jan-June PDSI

```{r, echo = FALSE, message = FALSE}

ch_up_puerco_wint_jj_plot <- ch_up_puerco_wint_jj %>% 
  mutate(mean_swe = mean_swe/10) %>% 
  select(waterYear.x, mean_swe, mean_pdsi) %>% 
  melt(id.vars = "waterYear.x") %>% 
  rename(metric = variable) %>% 
  ggplot(aes(x = waterYear.x, y = value, fill = metric)) +
  geom_col(position = "dodge") +
  theme_classic()

ch_up_puerco_wint_jj_plot

```

- 6/8 negative swe followed by negative pdsi
- 1/8 positive swe followed by positive pdsi

### Compare SWE of Chuska to PDSI chapters of interest

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# read in Tsaile chapter PDSI

tsaile_pdsi <- read_csv("Chapters/tsaile_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

tsaile_pdsi_mnth <- calc_pdsi_metrics(tsaile_pdsi, "month", mean)  
  plot_pdsi(tsaile_pdsi_mnth, title = "Tsaile")

tsaile_pdsi_lm <- lm(tsaile_pdsi_mnth$pdsi ~ tsaile_pdsi_mnth$date)
summary(tsaile_pdsi_lm)
```


#### Chuska SWE Anomalies vs Tsaile chapter PDSI
```{r, echo=FALSE, warning=FALSE, message = FALSE}
# graph chuska swe anomalies and tsaile pdsi:
ch_tsaile <- combine_swe_pdsi(tsaile_pdsi, ch_swe_anom)
plot_pdsi_swe(ch_tsaile, title = "Tsaile")

```

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# look at correlations
ch_tsaile_sp_esu <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(3,4), tsaile_pdsi_mnth, pdsi_mnths = c(5,6))
plot_cor_swe_pdsi(ch_tsaile_sp_esu, title = "Spring SWE and Tsaile early Summer PDSI")

```

#### Chuska Average SWE cold season vs PDSI of AMJ of Tsaile 
```{r, echo=FALSE, warning=FALSE, message = FALSE}
# what the anomaly ratio from difference looks like for Chuska
ggplot(ch_swe_anom_wint, aes(x = waterYear, y = anomaly_perc)) +
  geom_col() +
  theme_classic()


```

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# chuska winter vs AMJ of tsaile
ch_tsaile_wint_amj <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                     tsaile_pdsi_mnth, pdsi_mnths = c(4,5,6))
plot_cor_swe_pdsi(ch_tsaile_wint_amj, title = "Winter SWE and Tsaile AMJ PDSI")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

ch_tsail_amj_plot <- ch_tsaile_wint_amj %>% 
  mutate(mean_swe = mean_swe/10) %>% 
  select(waterYear.x, mean_swe, mean_pdsi) %>% 
  melt(id.vars = "waterYear.x") %>% 
  rename(metric = variable) %>% 
  ggplot(aes(x = waterYear.x, y = value, fill = metric)) +
  geom_col(position = "dodge") +
  theme_classic()

ch_tsail_amj_plot
```


```{r, echo = FALSE, warning=FALSE, message = FALSE}
# chuska winter vs A of tsaile
ch_tsaile_wint_a <- compare_swe_pdsi(ch_swe_anom, swe_mnths = c(11,12,1,2,3,4), 
                                     tsaile_pdsi_mnth, pdsi_mnths = c(4))
plot_cor_swe_pdsi(ch_tsaile_wint_amj, title = "Winter SWE and Tsaile April PDSI")

```

