---
title: "PDSI_final"
author: "AnnaClaire Marley"
date: "8/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
source("Functions/util.R")
source('Functions/graphing.R')
```

Finalized version of the PDSI markdwon

```{r, include = FALSE, message = FALSE}
# read in NN whole PDSI
nn_pdsi <- read_csv("NN_whole/nn_pdsi_79-19.csv") %>% 
  rename(date = DateTime,
         pdsi = `Palmer Drought Severity Index (PDSI) (gridMET) at 4005, 1979-01-01 to 2019-06-09`) %>% 
  add_water_year()
```

```{r, include = FALSE, warning=FALSE, message = FALSE}
# read in the agency pdsi 
chinle_pdsi <- read_csv("Agencies/chinle_pdsi.csv") 
east_navajo_pdsi <- read_csv("Agencies/east_navajo_pdsi.csv")
west_navajo_pdsi <- read_csv("Agencies/west_navajo_pdsi.csv")
shiprock_pdsi <- read_csv("Agencies/shiprock_pdsi.csv")
fort_defiance_pdsi <- read_csv("Agencies/fort_defiance_pdsi.csv")
```

```{r, include = FALSE}
### read in all the watershed pdsi###
chaco_pdsi <- read_csv("watersheds/chaco_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

chinle_pdsi_ws <- read_csv("watersheds/chinle_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

up_puerco_pdsi <- read_csv("watersheds/upper_puerco_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

mid_sj_pdsi <- read_csv("watersheds/middle_san_juan_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()
```

```{r, include = FALSE}
# read in mt swe
mtns_swe <- read_csv("NN_whole/nn_mtns_swe_total.csv")

# calculate winter anomaly
# methodology:
  # first calculate monthly mean swe, then take the mean for each water year
  # Need to make sure this method matches pdsi
mtns_swe_mnth_mean <- mtns_swe %>% 
  filter(month(date) %in% c(1,2,3,4,11,12)) %>% 
  rename(swe_mm = sum_mts_swe_m3) %>% 
  add_water_year() %>% 
  calc_swe_metrics("month", mean) 

mtns_swe_wint_anomaly <- mtns_swe_mnth_mean %>% 
  group_by(waterYear) %>% 
  summarize(mean_swe = mean(swe_mm)) %>% 
  mutate(anomaly = (mean_swe - mean(mean_swe))) %>% 
  mutate(anomaly_perc = anomaly/mean(mean_swe))
```


## Whole Navajo Nation
```{r, echo = FALSE, warning=FALSE}
# monthly pdsi values
nn_mnth_pdsi <- calc_pdsi_metrics(nn_pdsi, "month", mean)

plot_pdsi(nn_mnth_pdsi, title = "Navajo Nation")

nn_mnth_pdsi <- nn_mnth_pdsi %>% 
  filter(pdsi != "NA") 
  
nn_mnth_pdsi_lm <- lm(nn_mnth_pdsi$pdsi ~ nn_mnth_pdsi$date)
summary(nn_mnth_pdsi_lm)
```

- 1980s appeared to be more wet and then starting in 2000 there is a tendency to be in drought

```{r, echo = FALSE}
# let's look at the correlations between SPI and mtns swe for April-July SPI first
nn_mts_pdsi_apr <- merge_swe_pdsi(nn_mnth_pdsi, df_swe = mtns_swe_wint_anomaly, months = c(4)) 

nn_mts_pdsi_apr_plot <- plot_cor_swe_pdsi2(nn_mts_pdsi_apr,
                                          title = "Winter SWE Anomaly & April PDSI")

nn_mts_pdsi_may <- merge_swe_pdsi(nn_mnth_pdsi, df_swe = mtns_swe_wint_anomaly, months = c(5)) 
nn_mts_pdsi_may_plot <- plot_cor_swe_pdsi2(nn_mts_pdsi_may,
                                          title = "Winter SWE Anomaly & May PDSI")

nn_mts_pdsi_june <- merge_swe_pdsi(nn_mnth_pdsi, df_swe = mtns_swe_wint_anomaly, months = c(6)) 
nn_mts_pdsi_june_plot <- plot_cor_swe_pdsi2(nn_mts_pdsi_june,
                                          title = "Winter SWE Anomaly & June PDSI")

nn_mts_pdsi_july <- merge_swe_pdsi(nn_mnth_pdsi, df_swe = mtns_swe_wint_anomaly, months = c(7)) 
nn_mts_pdsi_july_plot <- plot_cor_swe_pdsi2(nn_mts_pdsi_july,
                                          title = "Winter SWE Anomaly & July PDSI")

```

```{r, echo=FALSE, fig.width=7}
grid.arrange(nn_mts_pdsi_apr_plot, nn_mts_pdsi_may_plot, 
             nn_mts_pdsi_june_plot, nn_mts_pdsi_july_plot, ncol = 2)
```

- High elevation winter swe anomaly is most correlated to Navajo Nation April PDSI 
  - the statistical significance isnt great for any of the correlations though

```{r, echo = FALSE}
# swe and pdsi for April pdsi

plot_swe_pdsi(nn_mnth_pdsi, month = c(4), swe_df = mtns_swe_wint_anomaly, region_time = "Navajo Nation April")

nn_mts_correct_apr <- swe_pdsi_same(nn_mts_pdsi_apr, type = "total")
nn_mts_positive_apr <- swe_pdsi_same(nn_mts_pdsi_apr, type = "positive")
nn_mts_negative_apr <- swe_pdsi_same(nn_mts_pdsi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(nn_mts_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r nn_mts_negative_apr*100`%
- *April* total same sign: `r nn_mts_correct_apr*100`%

## Agencies

```{r, include = FALSE, warning=FALSE}
# take the average yearly pdsi value 
chinle_an_av <- calc_pdsi_metrics(chinle_pdsi, "year", mean)
east_navajo_an_av <- calc_pdsi_metrics(east_navajo_pdsi, "year", mean)
west_navajo_an_av <- calc_pdsi_metrics(west_navajo_pdsi, "year", mean)
shiprock_an_av <- calc_pdsi_metrics(shiprock_pdsi, "year", mean)
fort_defiance_an_av <- calc_pdsi_metrics(fort_defiance_pdsi, "year", mean)

# take the monthly mean of the yearly data
chinle_mnth_av <- calc_pdsi_metrics(chinle_pdsi, "month", mean)
east_navajo_mnth_av <- calc_pdsi_metrics(east_navajo_pdsi, "month", mean)
west_navajo_mnth_av <- calc_pdsi_metrics(west_navajo_pdsi, "month", mean)
shiprock_mnth_av <- calc_pdsi_metrics(shiprock_pdsi, "month", mean)
fort_defiance_mnth_av <- calc_pdsi_metrics(fort_defiance_pdsi, "month", mean)
```

#### Correlation of PDSI monthly averages for the Agencies
```{r, echo = FALSE, warning=FALSE, message=FALSE}
## look to see how correlated the PDSI is between regions
pdsi_agencies <- chinle_mnth_av %>% 
  rename(chinle = pdsi) %>% 
  left_join(east_navajo_mnth_av, by = "date") %>% 
  rename(east_navajo = pdsi) %>% 
  left_join(west_navajo_mnth_av, by = "date") %>% 
  rename(west_navajo = pdsi) %>% 
  left_join(shiprock_mnth_av, by = "date") %>% 
  rename(shiprock = pdsi) %>% 
  left_join(fort_defiance_mnth_av, by = "date") %>% 
  rename(fort_defiance = pdsi) %>% 
  select(date, chinle, east_navajo, west_navajo, shiprock, fort_defiance) 
  
pdsi_cor_df<- pdsi_agencies %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "date")

# correlation matrix
cor_pdsi <- cor(pdsi_cor_df)
corrplot(cor_pdsi, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)

```

### Correlation between high elevation winter swe anomaly and the monthly average PDSI of the agencies

#### April SPI & All Mts Winter Swe Anomaly
```{r, echo=FALSE, warning=FALSE, message = FALSE}
## see how mts winter swe anomaly correlates with the pdsi of the agencies
agencies_v_mts <- pdsi_agencies %>% 
  filter(month(date) == 4) %>% 
  add_water_year() %>% 
  select(-date) %>% 
  left_join(mtns_swe_wint_anomaly, by = "waterYear") %>% 
  select(-waterYear, -mean_swe, -anomaly) %>% 
  rename(mts_swe_anom = anomaly_perc)

cor_pdsi_mts <- cor(agencies_v_mts)
corrplot(cor_pdsi_mts, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)
```

- Shiprock PDSI seems to be most correlated high elevation winter swe anomaly

### Correlation between Chuska winter swe anomaly and the monthly average PDSI of the agencies

```{r, echo = FALSE, message=FALSE}
# need to read in Chuska swe data and group data
chuska_swe <- read_csv("winter_wy/ch_swe_wy_03_19.csv")
chuska_swe_mnth_av <- calc_swe_metrics(chuska_swe, "month", mean)
chuska_mnth_anom <- calc_month_anom(chuska_swe_mnth_av) %>% 
  arrange(date)

# chuska winter swe anomaly
chuska_swe_wint_anom <- chuska_swe_mnth_av %>% 
  group_by(waterYear) %>% 
  summarize(mean_swe = mean(swe_mm)) %>% 
  ungroup() %>% 
  mutate(anomaly = (mean_swe - mean(mean_swe))) %>% 
  mutate(anomaly_perc = anomaly/mean(mean_swe))


```

#### April SPI and Chuska SWE Winter Anomaly
```{r, echo=FALSE, warning=FALSE, message = FALSE}
## see how chuska correlates with the pdsi of the agencies
agencies_v_chuska <- pdsi_agencies %>% 
  filter(month(date) == 4) %>% 
  add_water_year() %>% 
  select(-date) %>% 
  left_join(chuska_swe_wint_anom, by = "waterYear") %>% 
  select(-waterYear, -mean_swe, -anomaly) %>% 
  rename(chuska_anom = anomaly_perc)

cor_pdsi_chuska <- cor(agencies_v_chuska)
corrplot(cor_pdsi_chuska, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)
```

- Shiprocks April PDSI is most correlated to CHuska winter swe anomaly

## Watersheds

#### Watershed PDSI timeseries (1981-present)
```{r, echo = FALSE, warning=FALSE}
# visualize PDSI time series

chaco_pdsi_mnth <- calc_pdsi_metrics(chaco_pdsi, "month", mean) 
chaco_pdsi_mnth_plot <- plot_pdsi(chaco_pdsi_mnth, title = "Chaco")

chinle_pdsi_mnth_ws <- calc_pdsi_metrics(chinle_pdsi_ws, "month", mean) 
chinle_pdsi_mnth_ws_plot <-  plot_pdsi(chinle_pdsi_mnth_ws, title = "Chinle")

up_puerco_pdsi_mnth <- calc_pdsi_metrics(up_puerco_pdsi, "month", mean) 
up_puerco_pdsi_mnth_plot <-  plot_pdsi(up_puerco_pdsi_mnth, title = "Upper Puerco")

mid_sj_pdsi_mnth <- calc_pdsi_metrics(mid_sj_pdsi, "month", mean)
mid_sj_pdsi_mnth_plot <-  plot_pdsi(mid_sj_pdsi_mnth, title = "Middle San Juan")

```

```{r, echo = FALSE, fig.width = 7, warning = FALSE}
grid.arrange(chaco_pdsi_mnth_plot, chinle_pdsi_mnth_ws_plot, 
             up_puerco_pdsi_mnth_plot, mid_sj_pdsi_mnth_plot, ncol = 2)
```

- all of the watersheds show increasing drought towards the 21st century

#### Watershed PDSI vs Chuska winter swe anomaly

```{r, echo = FALSE}
# first join all the watersheds together
pdsi_watersheds <- chaco_pdsi_mnth %>% 
  rename(chaco = pdsi) %>% 
  left_join(chinle_pdsi_mnth_ws, by = "date") %>% 
  rename(chinle = pdsi) %>% 
  left_join(mid_sj_pdsi_mnth, by = "date") %>% 
  rename(middle_san_juan = pdsi) %>% 
  left_join(up_puerco_pdsi_mnth, by = "date") %>% 
  rename(upper_puerco = pdsi) %>% 
  select(date, chaco, chinle, middle_san_juan, upper_puerco)

```

##### April PDSI 
```{r, echo = FALSE}
# join chuska winter anomaly 
watersheds_v_chuska_pdsi <- pdsi_watersheds %>% 
  filter(month(date) == 4) %>% 
  add_water_year() %>% 
  filter(waterYear >= 2004) %>% 
  select(-date) %>% 
  left_join(chuska_swe_wint_anom, by = "waterYear") %>% 
  rename(chuska_swe_anom = anomaly_perc) %>% 
  select(-waterYear, -mean_swe, -anomaly)

cor_pdsi_ch <- cor(watersheds_v_chuska_pdsi)
corrplot(cor_pdsi_ch, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)


```

- Middle San Juan April PDSI has the highest correlation to Chuska winter swe anomaly

##### Middle San Juan April PDSI andC huska winter swe anomaly
```{r, echo = FALSE}
# swe and pdsi for April pdsi

plot_swe_pdsi(mid_sj_pdsi_mnth, month = c(4), swe_df = chuska_swe_wint_anom, region_time = "Middle San Juan April")

mid_sj_pdsi_apr <- merge_swe_pdsi(mid_sj_pdsi_mnth, df_swe = chuska_swe_wint_anom, months = c(4)) 

midsj_ch_correct_apr <- swe_pdsi_same(mid_sj_pdsi_apr, type = "total")
midsj_ch_positive_apr <- swe_pdsi_same(mid_sj_pdsi_apr, type = "positive")
midsj_ch_negative_apr <- swe_pdsi_same(mid_sj_pdsi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(midsj_ch_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r midsj_ch_negative_apr*100`%
- *April* total correct: `r midsj_ch_correct_apr*100`%


##### May
```{r, echo = FALSE}
# join chuska winter anomaly 
watersheds_v_chuska_pdsi_may <- pdsi_watersheds %>% 
  filter(month(date) == 5) %>% 
  add_water_year() %>% 
  filter(waterYear >= 2004) %>% 
  select(-date) %>% 
  left_join(chuska_swe_wint_anom, by = "waterYear") %>% 
  rename(chuska_swe_anom = anomaly_perc) %>% 
  select(-waterYear, -mean_swe, -anomaly)

cor_pdsi_ch_may <- cor(watersheds_v_chuska_pdsi_may)
corrplot(cor_pdsi_ch_may, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)

```

- for May PDSI, middle san juan is still most highly correlated to winter swe anomaly
- Watersheds' PDSI seem to be most highly correlated to winter swe anomaly in April

### Chuska winter swe and watershed PDSI AMJ 

```{r, echo = FALSE}
# join chuska winter anomaly 
watersheds_v_chuska_pdsi_aj <- pdsi_watersheds %>% 
  filter(month(date) %in% c(4, 5, 6)) %>% 
  add_water_year() %>% 
  filter(waterYear >= 2004) %>% 
  select(-date) %>% 
  group_by(waterYear) %>% 
  summarize(chaco = mean(chaco),
            chinle = mean(chinle),
            middle_san_juan = mean(middle_san_juan),
            upper_puerco = mean(upper_puerco)) %>% 
  left_join(chuska_swe_wint_anom, by = "waterYear") %>% 
  rename(chuska_swe_anom = anomaly_perc) %>% 
  select(-waterYear, -mean_swe, -anomaly)

cor_pdsi_ch_aj <- cor(watersheds_v_chuska_pdsi_aj)
corrplot(cor_pdsi_ch_aj, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)


```

### Chuska winter swe and watershed PDSI December - May (6 month PDSI)

```{r, echo = FALSE}
# join chuska winter anomaly 
watersheds_v_chuska_pdsi_dm <- pdsi_watersheds %>% 
  filter(month(date) %in% c(12, 1, 2, 3, 4, 5)) %>% 
  add_water_year() %>% 
  filter(waterYear >= 2004) %>% 
  select(-date) %>% 
  group_by(waterYear) %>% 
  summarize(chaco = mean(chaco),
            chinle = mean(chinle),
            middle_san_juan = mean(middle_san_juan),
            upper_puerco = mean(upper_puerco)) %>% 
  left_join(chuska_swe_wint_anom, by = "waterYear") %>% 
  rename(chuska_swe_anom = anomaly_perc) %>% 
  select(-waterYear, -mean_swe, -anomaly)

cor_pdsi_ch_dm <- cor(watersheds_v_chuska_pdsi_dm)
corrplot(cor_pdsi_ch_dm, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)


```

- not really correlated at all



### Chuska winter swe and watershed PDSI January - June (6 month PDSI)

```{r, echo = FALSE}
# join chuska winter anomaly 
watersheds_v_chuska_pdsi_jj <- pdsi_watersheds %>% 
  filter(month(date) %in% c(1, 2, 3, 4, 5, 6)) %>% 
  add_water_year() %>% 
  filter(waterYear >= 2004) %>% 
  select(-date) %>% 
  group_by(waterYear) %>% 
  summarize(chaco = mean(chaco),
            chinle = mean(chinle),
            middle_san_juan = mean(middle_san_juan),
            upper_puerco = mean(upper_puerco)) %>% 
  left_join(chuska_swe_wint_anom, by = "waterYear") %>% 
  rename(chuska_swe_anom = anomaly_perc) %>% 
  select(-waterYear, -mean_swe, -anomaly)

cor_pdsi_ch_jj <- cor(watersheds_v_chuska_pdsi_jj)
corrplot(cor_pdsi_ch_jj, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)


```

- a bit more highly correlated than Dec-May

### Tsaile Chapter

```{r, echo = FALSE, warning=FALSE, message = FALSE}
# read in Tsaile chapter PDSI

tsaile_pdsi <- read_csv("Chapters/tsaile_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

tsaile_pdsi_mnth <- calc_pdsi_metrics(tsaile_pdsi, "month", mean)  
  plot_pdsi(tsaile_pdsi_mnth, title = "Tsaile")

tsaile_pdsi_lm <- lm(tsaile_pdsi_mnth$pdsi ~ tsaile_pdsi_mnth$date)
summary(tsaile_pdsi_lm)
```


```{r, echo = FALSE}
# let's look at the correlations between SPI and mtns swe for April-July SPI first
ts_ch_pdsi_apr <- merge_swe_pdsi(tsaile_pdsi_mnth, df_swe = chuska_swe_wint_anom, months = c(4)) 

ts_ch_pdsi_apr_plot <- plot_cor_swe_pdsi2(ts_ch_pdsi_apr,
                                          title = "Winter SWE Anomaly & April PDSI")

ts_ch_pdsi_may <- merge_swe_pdsi(tsaile_pdsi_mnth, df_swe = chuska_swe_wint_anom, months = c(5)) 
ts_ch_pdsi_may_plot <- plot_cor_swe_pdsi2(ts_ch_pdsi_may,
                                          title = "Winter SWE Anomaly & May PDSI")

ts_ch_pdsi_june <- merge_swe_pdsi(tsaile_pdsi_mnth, df_swe = chuska_swe_wint_anom, months = c(6)) 
ts_ch_pdsi_june_plot <- plot_cor_swe_pdsi2(ts_ch_pdsi_june,
                                          title = "Winter SWE Anomaly & June PDSI")

ts_ch_pdsi_july <- merge_swe_pdsi(tsaile_pdsi_mnth, df_swe = chuska_swe_wint_anom, months = c(7)) 
ts_ch_pdsi_july_plot <- plot_cor_swe_pdsi2(ts_ch_pdsi_july,
                                          title = "Winter SWE Anomaly & July PDSI")
```

```{r, echo=FALSE, fig.width=7}
grid.arrange(ts_ch_pdsi_apr_plot, ts_ch_pdsi_may_plot, 
             ts_ch_pdsi_june_plot, ts_ch_pdsi_july_plot, ncol = 2)
```

- Tsaile PDSI doesnt really ever seem to be correlated with Chuska winter swe anomaly 

##### Tsaile April PDSI andC huska winter swe anomaly
```{r, echo = FALSE}
# swe and pdsi for April pdsi

plot_swe_pdsi(tsaile_pdsi_mnth, month = c(4), swe_df = chuska_swe_wint_anom, region_time = "Tsaile April")

tsaile_pdsi_apr <- merge_swe_pdsi(tsaile_pdsi_mnth, df_swe = chuska_swe_wint_anom, months = c(4)) 

ts_ch_correct_apr <- swe_pdsi_same(tsaile_pdsi_apr, type = "total")
ts_ch_positive_apr <- swe_pdsi_same(tsaile_pdsi_apr, type = "positive")
ts_ch_negative_apr <- swe_pdsi_same(tsaile_pdsi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(ts_ch_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r ts_ch_negative_apr*100`%
- *April* total correct: `r ts_ch_correct_apr*100`%

But Tsaile PDSI is almost consistently always negative, so be careful about interpreting the effect of SWE here