---
title: "PRISM_temp"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
source("Functions/util.R")
source('Functions/graphing.R')
```

This Rmarkdown uses data extracted by Suzanne Goldstein from Google EArth Engine

data include:

- monthly mean temperature data from PRISM  
- anomalies for each mountain region of the Navajo Nation during the period 2002-2018 as compared to the 30-year normal for the period 1981-2010. 

```{r, echo = FALSE, include = FALSE}

temp_anom <- read_csv("NN_whole/SG_mnthly_PRISM_anom_frm_30yr_norm_81_10.csv") %>% 
  clean_names()

temp <- read_csv("NN_whole/SG_monthly_PRISM.csv") %>% 
  clean_names()

temp_max <- read_csv("NN_whole/SG_mnthly_max_prism.csv") %>% 
  clean_names()

```

```{r, echo = FALSE, include = FALSE}
# chuska monthly winter swe
ch_wint_swe <- read_csv("winter_wy/ch_swe_wy_03_19.csv")

# chuska winter anomaly
ch_wint_anom <- read_csv("Chuska/ch_wint_swe_anom.csv") %>% 
  select(waterYear, anomaly_perc)

```

```{r, echo = FALSE}
# chuska monthly anomaly
ch_mnth_anom <- ch_wint_swe %>% 
  calc_swe_metrics("month", mean) %>% 
  calc_month_anom() %>% 
  arrange(date) 

```

### Mean winter temperature
```{r, echo = FALSE}
wint_temp_all <- temp %>% 
  filter(month(date) %in% c(11,12,1,2,3,4)) %>% 
  melt(id.vars = "date") %>% 
  add_water_year() %>% 
  group_by(waterYear, variable) %>% 
  summarize(mean_temp = mean(value)) 

ggplot(wint_temp_all, aes(x = waterYear, y = mean_temp)) +
  geom_line(aes(color = variable)) +
  labs(x = "Water Year",
       y = "Mean Winter Temperature (degrees C)") +
  scale_x_continuous(expand = c(0,0)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic()


```

## Chuska

### Average winter temperature
```{r, echo = FALSE}
# average winter temp
wint_temp_chuska <- temp %>% 
  filter(month(date) %in% c(11,12,1,2,3,4)) %>% 
  select(date, chuska) %>% 
  add_water_year() %>% 
  group_by(waterYear) %>% 
  summarize(mean_temp = mean(chuska)) 

ggplot(wint_temp_chuska, aes(x = waterYear, y = mean_temp)) +
  geom_line() +
  labs(x = "Water Year",
       y = "Mean Winter Temperature (degrees C)") +
  scale_x_continuous(expand = c(0,0)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic()

mean_temp_lm <- lm(mean_temp ~ waterYear, data = wint_temp_chuska)
summary(mean_temp_lm)

# slope is 0.05 and p value < 0.0001
```
- mean winter temperature has been increasing since 1980s
- slope of the line if 0.05 and p value < 0.0001

### Chuska winter swe anomaly and chuska winter temperature
```{r, echo = FALSE}
#chuska winter swe anomaly and mean winter temperature
ch_temp <- ch_wint_anom %>% 
  left_join(wint_temp_chuska, by = "waterYear") %>% 
  rename(swe_anom = anomaly_perc)

ggplot(ch_temp, aes(x = mean_temp, y = swe_anom)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Chuska Mean Winter Temperature (C)",
       y = "Chuska SWE Winter Anomaly")

cor.test(ch_temp$swe_anom, ch_temp$mean_temp)
```

- Winter mean temperature and winter swe anomaly are negative correlated (r = -4.1, but not significantly)

### monthly max
```{r}
# average of the winter monthly maximums
wint_av_max_chuska <- temp_max %>% 
  filter(month(date) %in% c(11,12,1,2,3,4)) %>% 
  select(date, chuska) %>% 
  add_water_year() %>% 
  group_by(waterYear) %>% 
  summarize(mean_temp_max = mean(chuska)) 

```

```{r, echo = FALSE}
ggplot(wint_av_max_chuska, aes(x = waterYear, y = mean_temp_max)) +
  geom_line() +
  labs(x = "Water Year",
       y = "Mean of Monthly Maximum Winter Temperatures (degrees C)") +
  scale_x_continuous(expand = c(0,0)) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic()
```

## Actual monthly SWE values compared to mean temperature chuska

```{r, echo = FALSE}
ch_swe_temp_av <- ch_wint_swe %>% 
  calc_swe_metrics("month", mean) %>% 
  left_join(temp, by = "date") %>% 
  select(date, swe_mm, chuska) %>% 
  mutate(month = month(date, label = TRUE))

ggplot(ch_swe_temp_av, aes(x = chuska, y = swe_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Mean Monthly Temperature (degrees C)",
    y = "Mean Monthly SWE (mm)"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic()

# correlation
cor.test(ch_swe_temp_av$swe_mm, ch_swe_temp_av$chuska)
# r = -0.27, p value = 0.008
```

### Relationship between Chuska winter swe and month average temperature
```{r, fig.width=7, echo = FALSE}
plot_month = function(df, name_month = "", title = "") {
    
    filter_month <- df %>% 
      filter(month == name_month)
  
    coef = cor.test(filter_month$swe_mm, filter_month$chuska)$estimate
    p_value = cor.test(filter_month$swe_mm, filter_month$chuska)$p.value
   
    plot <- filter_month %>% 
      ggplot(aes(x = chuska, y = swe_mm)) +
      geom_point() +
      geom_smooth(method = "lm", se = FALSE) +
      labs(
        x = "Mean Monthly Temperature (degrees C)",
        y = "Mean Monthly SWE (mm)",
        title = sprintf("%s", title),
        subtitle = sprintf("r = %s | p-value = %s", round(coef, 2), round(p_value,2))
      ) +
      scale_x_continuous(expand = c(0,0)) +
      theme_classic()
    return(plot)
}

jan <- plot_month(ch_swe_temp_av, name_month = "Jan", title = "January")
feb <- plot_month(ch_swe_temp_av, name_month = "Feb", title = "February")
mar <- plot_month(ch_swe_temp_av, name_month = "Mar", title = "March")
apr <- plot_month(ch_swe_temp_av, name_month = "Apr", title = "April")
nov <- plot_month(ch_swe_temp_av, name_month = "Nov", title = "November")
dec <- plot_month(ch_swe_temp_av, name_month = "Dec", title = "December")

grid.arrange(nov, dec, jan, feb, mar, apr, ncol = 3)

```


```{r, echo = FALSE}
cor.test(ch_swe_temp_av$chuska, ch_swe_temp_av$swe_mm)
summary(lm(ch_swe_temp_av$swe_mm ~ ch_swe_temp_av$chuska))

```

- Chuska monthly mean temperature and chuska mean monthly SWE have a negative correlation of 

```{r, echo = FALSE}
ch_swe_temp_av_max <- ch_wint_swe %>% 
  calc_swe_metrics("month", mean) %>% 
  left_join(temp_max, by = "date") %>% 
  select(date, swe_mm, chuska)

ggplot(ch_swe_temp_av_max, aes(x = chuska, y = swe_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Chuska Max Monthly Temp (C)",
    y = "Chuska Mean Monthly SWE (mm)"
  ) +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic()
```

```{r, echo = FALSE}

cor.test(ch_swe_temp_av_max$chuska, ch_swe_temp_av_max$swe_mm)

```


## Temperature anomaly from 1980-what? normals
```{r, echo = FALSE}
# average winter anomaly 
wint_anom_chuska <- temp_anom %>% 
  filter(month(date) %in% c(11,12,1,2,3,4)) %>% 
  select(date, chuska) %>% 
  add_water_year() %>% 
  group_by(waterYear) %>% 
  summarize(mean_anom = mean(chuska)) 

ggplot(wint_anom_chuska, aes(x = waterYear, y = mean_anom)) +
  geom_line() +
  theme_classic()
```

```{r, include=FALSE}
# read in Chuska Winter swe anomaly
ch_wint_anom <- read_csv("Chuska/ch_wint_swe_anom.csv") %>% 
  select(waterYear, anomaly_perc)
```

```{r, echo = FALSE}

ch_temp <- ch_wint_anom %>% 
  left_join(wint_anom_chuska, by = "waterYear") 


ggplot(ch_temp, aes(x = mean_anom, y = anomaly_perc)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Mean temperature anomaly",
    y = "Swe Anomaly"
  )


cor.test(ch_temp$anomaly_perc, ch_temp$mean_anom)
```
do i want to include temperature?


### Anomalies
```{r, echo = FALSE}

anom_graph <- temp_anom %>% 
  ggplot(aes(x = date, y = navajo_nation)) +
  geom_line(color = "red") +
  geom_smooth(method = "lm") +
  scale_x_date(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "Temperature Anomaly",
    title = "Navajo Nation PRISM Temperature Anomaly"
  ) +
  theme_classic() 

anom_graph

summary(lm(temp_anom$navajo_nation ~ temp_anom$date))

anom_graph_ch <- temp_anom %>% 
  ggplot(aes(x = date, y = chuska)) +
  geom_line(color = "red") +
  geom_smooth(method = "lm") +
  scale_x_date(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "Temperature Anomaly",
    title = "Chuska PRISM Temperature Anomaly"
  ) +
  theme_classic() 

anom_graph_ch

```

```{r, echo = FALSE}

# relationship between Chuska monthly temperature anomaly and chuska monthly swe anomaly
ch_temp_swe_anom <- temp_anom %>% 
  filter(year(date) >= 2003) %>% 
  filter(month(date) %in% c(11,12,1,2,3,4)) %>% 
  select(date, chuska) %>% 
  left_join(ch_mnth_anom, by = "date") %>% 
  tail(-4) %>% 
  select(date, waterYear, chuska, anomaly_perc) %>% 
  rename(temp_anom = chuska) %>% 
  mutate(swe_anom = anomaly_perc/100) %>% 
  select(-anomaly_perc)

cor.test(ch_temp_swe_anom$temp_anom, ch_temp_swe_anom$swe_anom)

```

```{r}
# visualize it
ggplot(ch_temp_swe_anom, aes(x = temp_anom, swe_anom)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

