---
title: "Raw data cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
# packages you need to clean the data
library(tidyverse)
library(janitor)
library(lubridate)
library(reshape2)
source("Functions/util.R")
```


# Snodas SWE Raw data

```{r, include = FALSE}
# read in the data:
carrizo <- read_csv("Carrizo/carrizo_swe_03_19.csv") 
chuska <- read_csv("Chuska/chuska_swe_03_19.csv")
black_mesa <- read_csv("Black Mesa/BM_swe_03_19.csv")
```

```{r, include = FALSE}
# Extract winter data
car_wint_cal <- extract_winter_months(carrizo)
ch_wint_cal <- extract_winter_months(chuska)
bm_wint_cal <- extract_winter_months(black_mesa)

# Adds water year. Renames column date to Date
car_wint_wy <- add_water_year(car_wint_cal)
ch_wint_wy <- add_water_year(ch_wint_cal)
bm_wint_wy <- add_water_year(bm_wint_cal)
```

```{r, include=FALSE}
# read in data
mt_powell_swe <- read_chunks("mt_powell", fill_pattern = ".*_swe_.*", variable = "swe_mm")

defiance_plat_swe <- read_chunks("defiance_plateau", fill_pattern = ".*_swe_.*", variable = "swe_mm")

navajo_mt_swe <- read_chunks("navajo_mtn", fill_pattern = ".*_snodas_swe_.*", variable = "swe_mm")

```

```{r, include = FALSE}
# extract winter months and add water year

mp_swe_wy <- month_ts(mt_powell_swe, c(11, 12, 1, 2, 3, 4)) %>%
            add_water_year()

dp_swe_wy <- month_ts(defiance_plat_swe, c(11, 12, 1, 2, 3, 4)) %>%
            add_water_year()

nm_swe_wy <- month_ts(navajo_mt_swe, c(11, 12, 1, 2, 3, 4)) %>%
            add_water_year()

```

```{r}
# now write the winter water year data as csv files

write.csv(bm_wint_wy, "Black Mesa/bm_swe_wint_wy_03_19.csv", row.names = FALSE)

write.csv(ch_wint_wy, "Chuska/ch_swe_wint_wy_03_19.csv", row.names = FALSE)

write.csv(car_wint_wy, "Carrizo/car_swe_wint_wy_03_19.csv", row.names = FALSE)

write.csv(mp_swe_wy, "mt_powell/mp_swe_wint_wy_03_19.csv", row.names = FALSE)

write.csv(dp_swe_wy, "defiance_plateau/dp_swe_wint_wy_03_19.csv", row.names = FALSE)

write.csv(nm_swe_wy, "navajo_mtn/nm_swe_wint_wy_03_19.csv", row.names = FALSE)

```

# Snodas snow Depth raw data

```{r, include = FALSE}
# clean into usable files from climate engine

# CARRIZO
car_snodas_depth <- read_chunks("Carrizo")

# CHUSKA
ch_snodas_depth <- read_chunks("Chuska")

# BLACK MESA
bm_snodas_depth <- read_chunks("Black Mesa")

# Navajo Mt
nm_snodas_depth <- read_chunks("navajo_mtn", fill_pattern = ".*_snodas_dpth_.*")

# Mt Powell

mp_snodas_depth <- read_chunks("mt_powell")

# Defiance Plateau
dp_snodas_depth <- read_chunks("defiance_plateau")

```

```{r, include = FALSE}
# extract winter months & add water year
car_depth_wy <- month_ts(car_snodas_depth, c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

ch_depth_wy <- month_ts(ch_snodas_depth, c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

bm_depth_wy <- month_ts(bm_snodas_depth, c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

nm_depth_wy <- month_ts(nm_snodas_depth, c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

mp_depth_wy <- month_ts(mp_snodas_depth, c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

dp_depth_wy <- month_ts(dp_snodas_depth, c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

```

```{r}
# now write the winter water year data as csv files
write.csv(bm_depth_wy, "Black Mesa/bm_dpth_wint_wy_03_19.csv", row.names = FALSE)

write.csv(ch_depth_wy, "Chuska/ch_dpth_wint_wy_03_19.csv", row.names = FALSE)

write.csv(car_depth_wy, "Carrizo/car_dpth_wint_wy_03_19.csv", row.names = FALSE)

write.csv(nm_depth_wy, "navajo_mtn/nm_dpth_wint_wy_03_19.csv", row.names = FALSE)

write.csv(mp_depth_wy, "mt_powell/mp_dpth_wint_wy_03_19.csv", row.names = FALSE)

write.csv(dp_depth_wy, "defiance_plateau/dp_dpth_wint_wy_03_19.csv", row.names = FALSE)

```


```{r}
## Navajo nation whole reservation
NN_swe <- read_chunks("NN_whole", fill_pattern = ".*_swe_.*", variable = "swe_mm") %>% 
  month_ts(c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year() %>% 
  rename(NN_swe_mm = swe_mm)

NN_depth <- read_chunks("NN_whole", fill_pattern = ".*_depth_.*", variable = "depth_mm") %>% 
  month_ts(c(11, 12, 1, 2, 3, 4)) %>%
  add_water_year()

```

Figure out the daily contributions of all mt range snow to navajo nation
This doesn't work actually because swe data is for 1km2 so the error is too big to make doing this make sense
```{r}
# combine daily values of swe of all regions into one dataframe

# read in swe data
car_swe <- read_csv("winter_wy/car_swe_wy_03_19.csv")
ch_swe <- read_csv("winter_wy/ch_swe_wy_03_19.csv")
bm_swe <- read_csv("winter_wy/bm_swe_wy_03_19.csv")
mp_swe <- read_csv("winter_wy/mp_swe_wy_03_19.csv")
nm_swe <- read_csv("winter_wy/nm_swe_wy_03_19.csv")
dp_swe <- read_csv("winter_wy/dp_swe_wy_03_19.csv")

# merge into same dataframe
all_swe <- car_swe %>% 
  rename(carrizo = swe_mm) %>% 
  merge(ch_swe, by = "date") %>% 
  rename(chuska = swe_mm) %>% 
  merge(bm_swe, by = "date") %>% 
  rename(black_mesa = swe_mm) %>% 
  select(-waterYear, -waterYear.x, -waterYear.y) %>% 
  merge(nm_swe, by = "date") %>% 
  rename(navajo_mt = swe_mm) %>% 
  merge(mp_swe, by = "date") %>% 
  rename(mt_powell = swe_mm) %>% 
  merge(dp_swe, by = "date") %>% 
  rename(defiance_plateau = swe_mm) %>% 
  select(-waterYear, -waterYear.x, -waterYear.y) 

# now I need to multiply each region's swe values by their area. Right now it's just the average swe value
# multiplying by the area will give me an approximate total swe value

# store all the areas as variables
carrizo_area = 339 #km2
chuska_area = 4581 #km2
bm_area = 1413 #km2
defiance_area = 1106 #km2
mt_powell_area = 77 #km2
nm_area = 96 #km2
nn_area = 67636 #km2


all_swe_total <- all_swe %>% 
  mutate(car_total = carrizo*carrizo_area,
         ch_total = chuska*chuska_area,
         bm_total = black_mesa*bm_area,
         dp_total = defiance_plateau*defiance_area,
         nm_total = navajo_mt*nm_area,
         mp_total = mt_powell*mt_powell_area) %>% 
  select(date, car_total, ch_total, bm_total, dp_total, nm_total, mp_total) %>% 
  mutate(mt_range_sum_total = rowSums(.[, names(.) != "date"])) # add across rows to get the sum of total swe in mts
 
# write this as a csv
  write.csv(all_swe_total, "winter_wy/all_swe_total.csv", row.names = FALSE)

# merge this data with the whole NN nation daily swe data and create a contribution column
NN_mt_range <- NN_swe %>% 
  mutate(NN_total = NN_swe_mm*nn_area) %>% 
  merge(all_swe_total, by = "date") %>% 
  select(date, waterYear, mt_range_sum_total, NN_total) %>% 
  mutate(contribution = mt_range_sum_total/NN_total) 

plot(NN_mt_range$date, NN_mt_range$contribution, type = "l")


# monthly average all regions swe
all_swe_sum_month_av <- all_swe_sum %>% 
  group_by(date = floor_date(date, "month"))  %>% 
  summarize(mt_ranges_mean_swe = mean(sum)) 

# now combine with NN nation
NN_all_swe_month_av <- calc_swe_metrics(NN_swe, "month", mean) %>% 
  merge(all_swe_sum_month_av, by = "date") %>% 
  mutate(contribution = mt_ranges_mean_swe/swe_mm)
  

```

- swe_mm of each mt range * it's area = total amount of swe
- then add them together to get total amount of water
- divide that number by the total amount of swe for nn 


## Palmer Drought Severity Index


```{r}
# Agencies
alamo_pdsi <- read_csv("data_download/data/ALAMO_pdsi_2003-09-01_2019-07-17.csv") %>% 
             clean_names() %>% 
            rename(pdsi = palmer_drounght_severity_index)

chinle_pdsi <- read_csv("data_download/data/CHINLE_pdsi_2003-09-01_2019-07-17.csv") %>% 
    clean_names() %>% 
    rename(pdsi = palmer_drounght_severity_index) %>% 
    add_water_year() 

east_navajo_pdsi <- read_csv("data_download/data/EASTERN NAVAJO_pdsi_2003-09-01_2019-07-17.csv") %>% 
             clean_names() %>% 
            rename(pdsi = palmer_drounght_severity_index) %>% 
             add_water_year() 
  
fort_defiance_pdsi <- read_csv("data_download/data/FORT DEFIANCE_pdsi_2003-09-01_2019-07-17.csv") %>% 
             clean_names() %>% 
            rename(pdsi = palmer_drounght_severity_index)%>% 
             add_water_year() 

shiprock_pdsi <- read_csv("data_download/data/SHIPROCK_pdsi_2003-09-01_2019-07-17.csv") %>% 
             clean_names() %>% 
            rename(pdsi = palmer_drounght_severity_index)%>% 
             add_water_year() 

west_navajo_pdsi <- read_csv("data_download/data/WESTERN NAVAJO_pdsi_2003-09-01_2019-07-17.csv") %>% 
             clean_names() %>% 
            rename(pdsi = palmer_drounght_severity_index)%>% 
             add_water_year() 

# write them as CSV as save in agency folder

write.csv(alamo_pdsi, "Agencies/alamo_pdsi.csv", row.names = FALSE)
write.csv(chinle_pdsi, "Agencies/chinle_pdsi.csv", row.names = FALSE)
write.csv(east_navajo_pdsi, "Agencies/east_navajo_pdsi.csv", row.names = FALSE)
write.csv(fort_defiance_pdsi, "Agencies/fort_defiance_pdsi.csv", row.names = FALSE)
write.csv(shiprock_pdsi, "Agencies/shiprock_pdsi.csv", row.names = FALSE)
write.csv(west_navajo_pdsi, "Agencies/west_navajo_pdsi.csv", row.names = FALSE)

```

# put the chinle watershed precip into one files
```{r}

chinle_precip <- read_chunks("data_download/data", fill_pattern = "Chinle_pr_.*", variable = "precipitation")

write.csv(chinle_precip, "watersheds/chinle_precip.csv", row.names = FALSE)

```



