---
title: "Snow Depth Analyses"
author: "AnnaClaire Marley"
date: "7/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
library(factoextra)
library(forcats)
source("Functions/util.R")
source('Functions/graphing.R')
```

```{r, include = FALSE}
# read in csv files - they are already processed as winter months and have a water year column. refer to "Raw_data_cleaning" for code 

car_depth_wy <- read_csv("Carrizo/car_dpth_wint_wy_03_19.csv") # Carrizo

ch_depth_wy <- read_csv("Chuska/ch_dpth_wint_wy_03_19.csv") # Chuska

bm_depth_wy <- read_csv("Black Mesa/bm_dpth_wint_wy_03_19.csv") # Black Mesa

nm_depth_wy <- read_csv("navajo_mtn/nm_dpth_wint_wy_03_19.csv") # Navajo Mt

mp_depth_wy <- read_csv("mt_powell/mp_dpth_wint_wy_03_19.csv") # Mt Powell
  
dp_depth_wy <- read_csv("defiance_plateau/dp_dpth_wint_wy_03_19.csv") # Defiance Plateau
```


### Daily Snow Depth Variability for Each Region (November - April Water Year)

- 2004-2007, 2012, and 2018 were especially dry years across all regions
  - 2004 & 2005 experienced almost no snow
- Carrizo has the smallest snow depths and Chuska has the higest

```{r, fig.width=20,fig.height=5, echo= FALSE}
carr_dpth_daily <- graph_with_wateryear(car_depth_wy, "Carrizo", variable = "depth_mm", ylab = "Snow depth (mm)")
carr_dpth_daily
```

```{r, fig.width=20,fig.height=5, echo= FALSE}
ch_dpth_daily <- graph_with_wateryear(ch_depth_wy, "Chuska", variable = "depth_mm", ylab = "Snow depth (mm)")
ch_dpth_daily
```

```{r, fig.width=20,fig.height=5, echo= FALSE}
bm_dpth_daily <- graph_with_wateryear(bm_depth_wy, "Black Mesa", variable = "depth_mm", ylab = "Snow depth (mm)")
bm_dpth_daily
```

### Annual Mean Snow Depth variability (November - April Water Year)

- Lowest average snow depth across all three regions = 2004-2007 WY and 2018
- Chuska on average always has the deepest snow

```{r, echo = FALSE, include = FALSE}
car_an_av <- graph_anwy_metrics(car_depth_wy, mean, title = "Carrizo")
ch_an_av <- graph_anwy_metrics(ch_depth_wy, mean, title = "Chuska")
bm_an_av <- graph_anwy_metrics(bm_depth_wy, mean, title = "Black Mesa")
nm_an_av
mp_an_av
dp_an_an

#graphs

car_an_av
ch_an_av
bm_an_av

```

```{r, echo = FALSE}
# graph annual average stacked on top of each other

# quick function to calculate annual averages
calc_an_region = function(df) {

  av = df %>% 
    group_by(waterYear)  %>% 
    summarize(depth_mm = mean(depth_mm))

return(av)
}

# make the annual average dataframes
car_an <- calc_an_region(car_depth_wy)
ch_an <- calc_an_region(ch_depth_wy)
bm_an <- calc_an_region(bm_depth_wy)
nm_an <- calc_an_region(nm_depth_wy)
mp_an <- calc_an_region(mp_depth_wy)
dp_an <- calc_an_region(dp_depth_wy)

# join together
an_av_all <- join_depth_months("Carrizo" = car_an, "Chuska" = ch_an, 
                               "Navajo Mt." = nm_an, "Black Mesa" = bm_an,
                               "Mt. Powell" = mp_an, "Defiance Plateau" = dp_an)


# graph it
an_av_depth_facet <- an_av_all %>% 
   ggplot(aes(x = waterYear, y = depth_mm)) +
  geom_col() +
  facet_wrap(~ factor(region, levels = c("Chuska", "Defiance Plateau",
                                         "Black Mesa", "Mt. Powell",
                                         "Carrizo", "Navajo Mt.")
                      
                      )) +
  scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Water Year",
    y = "Snow Depth (mm)"
  ) +
  theme_bw()


# graph it

an_av_depth_graph <- an_av_all %>% 
  ggplot(aes(x = waterYear, y = depth_mm, fill = region)) +
  geom_bar(stat = "identity", position = "identity") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_discrete(name = "Region", labels = c("Carrizo", "Chuska", "Black Mesa")) +
  labs(
    x = "Water Year",
    y = "Snow depth (mm)",
    region = "Region"
  ) +
  theme_classic()


```




## Monthly Mean Snow Depth variability (November - April Water Year)

```{r, include = FALSE}
# monthly mean
car_mn_av <- calc_depth_metrics(car_depth_wy, "month", mean)
ch_mn_av <- calc_depth_metrics(ch_depth_wy, "month", mean)
bm_mn_av <- calc_depth_metrics(bm_depth_wy, "month", mean)
nm_mn_av <- calc_depth_metrics(nm_depth_wy, "month", mean)
mp_mn_av <- calc_depth_metrics(mp_depth_wy, "month", mean)
dp_mn_av <- calc_depth_metrics(dp_depth_wy, "month", mean)

#monthly max
car_mn_max <- calc_depth_metrics(car_depth_wy, "month", max)
ch_mn_max <- calc_depth_metrics(ch_depth_wy, "month", max)
bm_mn_max <- calc_depth_metrics(bm_depth_wy, "month", max)
nm_mn_max <- calc_depth_metrics(nm_depth_wy, "month", max)
mp_mn_max <- calc_depth_metrics(mp_depth_wy, "month", max)
dp_mn_max <- calc_depth_metrics(dp_depth_wy, "month", max)
```

Carrizo: 

- Low years: 2004-2009, 2012, 2015, 2018

```{r,fig.width=20,fig.height=5, echo = FALSE}
# graph it
# carrizo 
car_mn_av_graph <- graph_with_wateryear(car_mn_av, "Carrizo", variable = "depth_mm", ylab = "Snow depth (mm)")
car_mn_av_graph

```

Chuska:

- low years: 2004-2007, 2014, 2018

```{r,fig.width=20,fig.height=5, echo = FALSE}
# graph it
# chuska
ch_mn_av_graph <- graph_with_wateryear(ch_mn_av, "Chuska", variable = "depth_mm", ylab = "Snow depth (mm)")
ch_mn_av_graph

```

Black Mesa:

- low years: 2004-2007, 2011-2012, 2014-2015, 2018


```{r,fig.width=20,fig.height=5, echo = FALSE}
# graph it
# BLack Mesa
bm_mn_av_graph <- graph_with_wateryear(bm_mn_av, "Black Mesa", variable = "depth_mm", ylab = "Snow depth (mm)")
bm_mn_av_graph
```


### Correlation of monthly mean snow depths
```{r, echo = FALSE}
# correlation matrix of monthly variability

# first make the dataframe
av_month_depth_all <- car_mn_av %>% 
  rename(carrizo = depth_mm) %>% 
  merge(ch_mn_av, by = "date") %>% 
  rename(chuska = depth_mm) %>% 
  merge(bm_mn_av, by = "date") %>% 
  rename(black_mesa = depth_mm) %>% 
  select(-waterYear, -waterYear.x, -waterYear.y) %>% 
  merge(nm_mn_av, by = "date") %>% 
  rename(navajo_mt = depth_mm) %>% 
  merge(mp_mn_av, by = "date") %>% 
  rename(mt_powell = depth_mm) %>% 
  merge(dp_mn_av, by = "date") %>% 
  rename(defiance_plateau = depth_mm) %>% 
  select(-waterYear, -waterYear.x, -waterYear.y) %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "date")


```

```{r, echo = FALSE}
# correlation matrix
cor <- cor(av_month_depth_all)
corrplot(cor, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)

```

- Least correlated are mt powell and navajo mt. This makes sense as they are the furthest sites from each other
- Next is defiance plateau and navajo mt which are the next farthest from each other
- This trend does not continue, because navajo mt and black mesa are the next least correlated. However, Black Mesa is the cosest mountain range to navajo mt.
  but black mesa appears to be much lower elevation than navajo mt, which may explain
- The higest correlation is between black mesa and defiance plateau. They are relatively close and appear to have similar elevations

## Anomalies

Difference from the monthly mean for each region


Carrizo:

- Anomalously dry years: 2004-2008, 2011-2012, 2018
  - most of 2009, 2014 and 2015 also mostly negatively anomalous
```{r, fig.width=20,fig.height=5,echo = FALSE}

# Carrizo
car_dpth_anom <- calc_mnth_depth_anom(car_mn_av)
plot_anomaly(car_dpth_anom)

```


```{r, include = FALSE}
# if you want to look closely at any one month use this code:

# carrizo december anomaly:

# plot_monthly_anomaly(car_dpth_anom, 12, title = "Carrizo December")


```


Chuska:

- Anomalously dry years: 2004-2007, 2015, 2018

```{r, fig.width=20,fig.height=5,echo = FALSE}
# chuska
ch_dpth_anom <- calc_mnth_depth_anom(ch_mn_av)
plot_anomaly(ch_dpth_anom)

```

Black Mesa:

- Anomalously dry years: 2004-2007, 2011, 2012, 2014, 2015, 2018


```{r, fig.width=20,fig.height=5,echo = FALSE}
# black mesa
bm_dpth_anom<- calc_mnth_depth_anom(bm_mn_av)
plot_anomaly(bm_dpth_anom)

```

### Weekly Mean Correlations between sites

```{r, include = FALSE}

car_wk_av <- calc_depth_metrics(car_depth_wy, "1 week", mean)
ch_wk_av <- calc_depth_metrics(ch_depth_wy, "1 week", mean)
bm_wk_av <- calc_depth_metrics(bm_depth_wy, "1 week", mean)

```

```{r, include = FALSE}
## Correlation tests to find r2 values

# carrizo vs chuska

car_v_ch_wk_av_cor <- cor.test(car_wk_av$depth_mm, ch_wk_av$depth_mm)$estimate # r value

car_v_ch_wk_av_r2 <- (car_v_ch_wk_av_cor)^2 # r squared

# carrizo vs black mesa
car_v_bm_wk_av_cor <- cor.test(car_wk_av$depth_mm, bm_wk_av$depth_mm)$estimate

car_v_bm_wk_av_r2 <- (car_v_bm_wk_av_cor)^2 # r squared


# black mesa vs chuska
bm_v_ch_wk_av_cor <- cor.test(bm_wk_av$depth_mm, ch_wk_av$depth_mm)$estimate

bm_v_ch_wk_av_r2 <- (bm_v_ch_wk_av_cor)^2

```

Pearson's R Squared values:

*Black Mesa and Carrizo most closely correlated even though Carrizo and the Chuskas are closer and have more similar elevations. Black Mesa and Chuska least closely correlated*

- Carrizo vs Chuska: `r round(car_v_ch_wk_av_r2, 2)`

- Carrizo vs Black Mesa: `r round(car_v_bm_wk_av_r2, 2)`

- Black Mesa vs Chuska: `r round(bm_v_ch_wk_av_r2, 2)`

### Average snowdepth values for each months at each site

```{r, echo = FALSE}
# november
car_nov_dpth <- month_ts(car_mn_av, 11)
ch_nov_dpth <- month_ts(ch_mn_av, 11)
bm_nov_dpth <- month_ts(bm_mn_av, 11)
nm_nov_dpth <- month_ts(nm_mn_av, 11)
dp_nov_dpth <- month_ts(dp_mn_av, 11) 
mp_nov_dpth <- month_ts(mp_mn_av, 11)

nov_dpth <- join_depth_months(carrizo = car_nov_dpth, chuska = ch_nov_dpth, black_mesa = bm_nov_dpth, defiance_plateau = dp_nov_dpth, mt_powell = mp_nov_dpth, navajo_mt = nm_nov_dpth)

nov_dpth_graph <- month_bplot_dpth(nov_dpth, title = "November")

```


```{r, error = TRUE, include = FALSE}
# PCA?
#nov_dpth_swe <- merge(nov_dpth, nov_swe, by = c("waterYear", "region"))
#pcs <- prcomp(nov_dpth_swe[3:4], scale = TRUE)
#summary(pcs)

#group <-  nov_dpth_swe$region

#ggbiplot(pcs, groups = group)

#fviz_pca_ind(pcs, col.ind = group)

```

   
```{r, echo = FALSE}
# December

car_dec_dpth <- month_ts(car_mn_av, 12)
ch_dec_dpth <- month_ts(ch_mn_av, 12)
bm_dec_dpth <- month_ts(bm_mn_av, 12)
nm_dec_dpth <- month_ts(nm_mn_av, 12)
dp_dec_dpth <- month_ts(dp_mn_av, 12) 
mp_dec_dpth <- month_ts(mp_mn_av, 12)

dec_dpth <- join_depth_months(carrizo = car_dec_dpth, chuska = ch_dec_dpth, black_mesa = bm_dec_dpth, defiance_plateau = dp_dec_dpth, mt_powell = mp_dec_dpth, navajo_mt = nm_dec_dpth)

dec_dpth_graph <- month_bplot_dpth(dec_dpth, title = "December") +
  scale_x_discrete(labels = c("Chuska", "Defiance Plateau", "Black Mesa", "Mt. Powell", "Carrizo", "Navajo Mt."))

```



```{r, echo = FALSE}
# January

jan_dpth <- join_depth_months(carrizo = month_ts(car_mn_av, 1), 
                              chuska = month_ts(ch_mn_av, 1),
                              black_mesa = month_ts(bm_mn_av, 1), 
                              defiance_plateau = month_ts(dp_mn_av, 1) , 
                              mt_powell = month_ts(mp_mn_av, 1), 
                              navajo_mt = month_ts(nm_mn_av, 1))

jan_dpth_graph <- month_bplot_dpth(jan_dpth, title = "January") +
  scale_x_discrete(labels = c("Chuska", "Defiance Plateau",  "Mt. Powell", "Black Mesa","Carrizo", "Navajo Mt."))

```



```{r, echo = FALSE}
# Februrary

feb_dpth <- join_depth_months(carrizo = month_ts(car_mn_av, 2), 
                              chuska = month_ts(ch_mn_av, 2),
                              black_mesa = month_ts(bm_mn_av, 2), 
                              defiance_plateau = month_ts(dp_mn_av, 2) , 
                              mt_powell = month_ts(mp_mn_av, 2), 
                              navajo_mt = month_ts(nm_mn_av, 2))

feb_dpth_graph <- month_bplot_dpth(feb_dpth, title = "February") +
  scale_x_discrete(labels = c("Chuska", "Defiance Plateau",  "Mt. Powell", "Carrizo", "Black Mesa", "Navajo Mt."))

```


```{r, echo = FALSE}
# March

mar_dpth <- join_depth_months(carrizo = month_ts(car_mn_av, 3), 
                              chuska = month_ts(ch_mn_av, 3),
                              black_mesa = month_ts(bm_mn_av, 3), 
                              defiance_plateau = month_ts(dp_mn_av, 3) , 
                              mt_powell = month_ts(mp_mn_av, 3), 
                              navajo_mt = month_ts(nm_mn_av, 3))

mar_dpth_graph <- month_bplot_dpth(mar_dpth, title = "March") +
  scale_x_discrete(labels = c("Chuska", "Defiance Plateau",  "Black Mesa", "Mt. Powell", "Carrizo",  "Navajo Mt."))

```


```{r, echo = FALSE}
# April

apr_dpth <- join_depth_months(carrizo = month_ts(car_mn_av, 4), 
                              chuska = month_ts(ch_mn_av, 4),
                              black_mesa = month_ts(bm_mn_av, 4), 
                              defiance_plateau = month_ts(dp_mn_av, 4) , 
                              mt_powell = month_ts(mp_mn_av, 4), 
                              navajo_mt = month_ts(nm_mn_av, 4))

apr_dpth_graph <- month_bplot_dpth(apr_dpth, title = "April") +
  scale_x_discrete(labels = c("Chuska", "Defiance Plateau",  "Black Mesa", "Navajo Mt.", "Carrizo", "Mt. Powell"))

```



```{r, echo = FALSE, fig.width=20, fig.height=5}

grid.arrange(nov_dpth_graph, dec_dpth_graph, jan_dpth_graph, feb_dpth_graph, mar_dpth_graph, apr_dpth_graph, ncol = 3)

```

- peak snow depth is usually January and sometimes February
- decreases drastically in february and then is mostly gone by april
- Navajo my usually has the shallowest snow depth

```{r, fig.width=15, echo = FALSE}

grid.arrange(dec_dpth_graph, jan_dpth_graph, feb_dpth_graph, mar_dpth_graph, ncol = 2)

```



```{r, echo = FALSE}
# Since max snow depth usually occurs in January, let's see how max snow depth is changing over time 

max_Depth <- join_depth_months(carrizo = month_ts(car_mn_max, 1), 
                              chuska = month_ts(ch_mn_max, 1),
                              black_mesa = month_ts(bm_mn_max, 1), 
                              defiance_plateau = month_ts(dp_mn_max, 1) , 
                              mt_powell = month_ts(mp_mn_max, 1), 
                              navajo_mt = month_ts(nm_mn_max, 1))

all_max_depth_graph <- max_Depth %>% 
  ggplot(aes(x = waterYear, y = depth_mm, group = region, color = region)) +
  geom_line() +
  theme_classic() 

```

