---
title: "SPI from CHRPS"
author: "AnnaClaire Marley"
date: "8/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
library(precintcon)
library(onewaytests)
library(car)
source("Functions/util.R")
source('Functions/graphing.R')
```

- **This Rmarkdown changes CHRPS daily precipitation (mm) to SPI**

```{r, echo = FALSE, message = FALSE}
# read in precipitation files
nn_chrps_precip <- read_csv("NN_whole/nn_chrps_precip_80-19.csv") %>% 
  arrange(date)

mid_sj_precip <- read_csv("watersheds/mid_sj_chirps.csv") %>% 
  arrange(date)

chaco_precip <- read_csv("watersheds/chaco_chirps.csv", 
                         col_names=c("date", "precipitation"), 
                         skip =1) %>% 
  arrange(date)

up_pue_precip <- read_csv("watersheds/up_pue_chirps.csv") %>% 
  arrange(date)

chinle_precip <- read_csv("watersheds/chinle_chirps.csv") %>% 
  arrange(date)

# Tsaile chapter chirps:
tsaile_precip <- read_csv("Chapters/tsaile_chirps.csv")

# for later use:
all_months = c(1,2,3,4,5,6,7,8,9,10,11,12)
```


```{r, echo = FALSE}
# read in all mountains winter swe anomaly
mtns_swe <- read_csv("NN_whole/nn_mtns_swe_total.csv")

# calculate winter anomaly
# methodology:
  # first calculate monthly mean swe, then take the mean for each water year
  # Need to make sure this method matches pdsi
mtns_swe_mnth_mean <- mtns_swe %>% 
  filter(month(date) %in% c(1,2,3,4,11,12)) %>% 
  rename(swe_mm = sum_mts_swe_m3) %>% 
  add_water_year() %>% 
  calc_swe_metrics("month", mean) 

mtns_swe_wint_anomaly <- mtns_swe_mnth_mean %>% 
  group_by(waterYear) %>% 
  summarize(mean_swe = mean(swe_mm)) %>% 
  mutate(anomaly = (mean_swe - mean(mean_swe))) %>% 
  mutate(anomaly_perc = anomaly/mean(mean_swe))
```

```{r, include=FALSE}
# read in Chuska Winter swe anomaly
ch_wint_anom <- read_csv("Chuska/ch_wint_swe_anom.csv") %>% 
  select(waterYear, anomaly_perc)
```

```{r, echo = FALSE}
# read in Day of year snowmelt
ch_snowmelt <- read_csv("Chuska/ch_doy_snowmelt.csv")
```


## Whole Navajo Nation
### Precipitation
```{r, echo = FALSE}
# monthly precipitation
nn_precip_mnth <- nn_chrps_precip %>% 
  group_by(date = floor_date(date, "month")) %>% 
  summarize(precipitation = mean(precipitation)) %>% 
  ungroup()

ggplot(nn_precip_mnth, aes(x = date, y = precipitation)) +
  geom_line() +
  scale_x_date(expand = c(0,0)) +
  labs(
    title = "Navajo Nation Monthly Mean Precipitation"
  ) +
  theme_classic()

```

```{r, echo = FALSE}
# make monthly precipitation anomaly based on 1981-2001 month averages
  precip_intra_month = nn_precip_mnth %>% 
    arrange(date) %>% 
    filter(year(date) <= 2001) %>% 
    mutate(month = month(date)) %>% 
    group_by(month) %>% 
    summarize(mean_precipitation = mean(precipitation))
  
  precip_intra_month_anomaly = nn_precip_mnth %>% 
    mutate(month = month(date)) %>% 
    merge(precip_intra_month, by = "month")%>%
    mutate(anomaly = precipitation - mean_precipitation) %>% 
    mutate(anomaly_ratio = (anomaly/mean_precipitation)) %>% 
    mutate(sign = ifelse(anomaly < 0, "negative", "positive") ) %>% 
    arrange(date)
    

```

```{r, echo = FALSE, fig.width=7}
# graph it
precip_anomaly_march <- precip_intra_month_anomaly %>% 
  filter(month(date) %in% c(3)) %>% 
  ggplot(aes(x = date, y = anomaly_ratio)) +
  geom_col(aes(fill = sign)) +
  labs(title = "March Anomaly") +
  scale_fill_manual(values = c("negative" = "red", "positive" = "dark green")) +
  theme_classic()

precip_anomaly_apr <- precip_intra_month_anomaly %>% 
  filter(month(date) %in% c(4)) %>% 
  ggplot(aes(x = date, y = anomaly_ratio)) +
  geom_col(aes(fill = sign)) +
  labs(title = "April Anomaly") +
  scale_fill_manual(values = c("negative" = "red", "positive" = "dark green")) +
  theme_classic()


grid.arrange(precip_anomaly_march, precip_anomaly_apr, ncol = 2)
```


#### 6 month SPI
```{r, echo = FALSE}
# change precipitation to SPI
nn_spi <- precip_to_spi(nn_chrps_precip)

# check something

```

```{r, echo = FALSE}
# look at it
plot_spi(nn_spi, month = all_months, title = "", years = "all") 

# the linear model info:
nn_spi_lm <- lm(nn_spi$spi ~ nn_spi$date)
summary(nn_spi_lm)
```


#### NN SPI vs PDSI
```{r, echo = FALSE}
# read in pdsi and get monthly mean
nn_pdsi <- read_csv("NN_whole/nn_pdsi_79-19.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year()

nn_pdsi_mnth <- nn_pdsi %>% 
  calc_pdsi_metrics("month", mean) %>% 
  rename(mean_pdsi = pdsi)

#merge with NN SPI
nn_pdsi_spi <- nn_spi %>% 
  select(date, spi) %>% 
  merge(nn_pdsi_mnth, by = "date") 

# plot their correlation
plot_cor_pdsi_spi(nn_pdsi_spi, title = "Navajo Nation")
```

- SPI and PDSI have a pretty high correlation

#### All mountains SWE vs NN SPI

```{r, echo = FALSE, fig.width=7}
# let's look at the correlations between SPI and mtns swe for April-July SPI first
plot_spi_swe_all(nn_spi, mtns_swe_wint_anomaly)

```

- April SPI and high elevation swe anomaly have strongest correlation

#### without 2010
```{r, echo = FALSE}

# let's look at the correlations between SPI and mtns swe for April-July SPI WITHOUT 2010
nn_mts_spi_apr_n10 <- merge_ch_spi(nn_spi, df_swe = mtns_swe_wint_anomaly, months = c(4)) %>% 
  filter(waterYear != 2010)

nn_mts_spi_apr_plot_n10 <- plot_cor_swe_spi(nn_mts_spi_apr_n10,
                                          title = "Winter SWE Anomaly & April SPI")

nn_mts_spi_may_n10 <- merge_ch_spi(nn_spi, df_swe = mtns_swe_wint_anomaly, months = c(5)) %>% 
  filter(waterYear != 2010)
nn_mts_spi_may_plot_n10 <- plot_cor_swe_spi(nn_mts_spi_may_n10,
                                          title = "Winter SWE Anomaly & May SPI")

nn_mts_spi_june_n10 <- merge_ch_spi(nn_spi, df_swe = mtns_swe_wint_anomaly, months = c(6)) %>% 
  filter(waterYear != 2010)
nn_mts_spi_june_plot_n10 <- plot_cor_swe_spi(nn_mts_spi_june_n10,
                                          title = "Winter SWE Anomaly & June SPI")

nn_mts_spi_july_n10 <- merge_ch_spi(nn_spi, df_swe = mtns_swe_wint_anomaly, months = c(7)) %>% 
  filter(waterYear != 2010)
nn_mts_spi_july_plot_n10 <- plot_cor_swe_spi(nn_mts_spi_july_n10,
                                          title = "Winter SWE Anomaly & July SPI")

```

```{r, echo=FALSE, message = FALSE, fig.width=7}

grid.arrange(nn_mts_spi_apr_plot_n10, nn_mts_spi_may_plot_n10, 
             nn_mts_spi_june_plot_n10, nn_mts_spi_july_plot_n10, ncol = 2)

```

- taking out 2010 actually just makes it worse

```{r, echo = FALSE}
nn_mts_spi_may <- merge_ch_spi(nn_spi, df_swe = mtns_swe_wint_anomaly, months = c(5))

# swe and spi following for MAY spi

plot_swe_spi(nn_spi, month = c(5), 
             swe_df = mtns_swe_wint_anomaly, region_time = "Navajo Nation May 6 month")

nn_mts_correct_may <- swe_spi_same(nn_mts_spi_may, type = "total")
nn_mts_positive_may <- swe_spi_same(nn_mts_spi_may, type = "positive")
nn_mts_negative_may <- swe_spi_same(nn_mts_spi_may, type = "negative")
```

- *May* positive swe & positive spi / positive swe: `r round(nn_mts_positive_may*100, 3)`%
- *May* negative swe & negative spi / negative swe: `r nn_mts_negative_may*100`%
- *May* total correct: `r nn_mts_correct_may*100`%

```{r, echo = FALSE}

plot_spi_drought_level(nn_spi, month = c(5), region_time = "Navajo Nation May 6 month")


```

```{r, echo = FALSE}
nn_mts_spi_apr <- merge_ch_spi(nn_spi, df_swe = mtns_swe_wint_anomaly, months = c(4))

# swe and spi following for April spi
plot_swe_spi(nn_spi, month = c(4), swe_df = mtns_swe_wint_anomaly, region_time = "Navajo Nation April 6 month")

nn_mts_correct_apr <- swe_spi_same(nn_mts_spi_apr, type = "total")
nn_mts_positive_apr <- swe_spi_same(nn_mts_spi_apr, type = "positive")
nn_mts_negative_apr <- swe_spi_same(nn_mts_spi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(nn_mts_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r nn_mts_negative_apr*100`%
- *April* total correct: `r nn_mts_correct_apr*100`%

```{r, echo = FALSE}

plot_spi_drought_level(nn_spi, month = c(4), region_time = "Navajo Nation April 6 month")

```

- every time in april there was a drought warning or a drought emergency, it was preceded by a negative anomaly winter swe


```{r, echo = FALSE, fig.width=7}
plot_spi_drought_level(nn_spi, month = c(4), 
                       region_time = "Navajo Nation April 6 month") 

plot_spi_drought_level(nn_spi, month = all_months, 
                       region_time = "Navajo Nation 6 month", years = "all") 
```

- drought emergency occured in 2018, with drought warnings in 2006 and 2012

```{r, echo = FALSE}
# what were the swe anomalies for years that had a drought warning/drought emergency in april?

nn_drought_times <- nn_mts_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  filter(drought == "warning" | drought == "emergency")
nn_drought_times


nn_swe_drought_apr <- nn_mts_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  ggplot(aes(x = waterYear, y = swe_anom,
             fill = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal"))
               )) +
  geom_col() +
  scale_fill_manual(values = c("#d13111", 
                                "#ffa600", 
                                "#f1d333", 
                                "#16990c"), 
                     labels = c("Drought Emergency", 
                                "Drought Warning", 
                                "Drought Alert", 
                                "Normal")) +
  labs(
    x = "Water Year",
    y = "SWE",
    fill = "Drought Severity",
    title = "Winter Swe Anomaly colored by April SPI drought warning"
  ) +
  theme_classic()

nn_swe_drought_apr
```


## Watersheds 

```{r, echo = FALSE}
# change to spi
mid_sj_spi <- precip_to_spi(mid_sj_precip)
chaco_spi <- precip_to_spi(chaco_precip)
chinle_spi <- precip_to_spi(chinle_precip)
up_pue_spi <- precip_to_spi(up_pue_precip)
```

```{r, include = FALSE}
# combine together to make a correlation matrix
watersheds_spi <- mid_sj_spi %>% 
  rename(middle_sj = spi) %>% 
  left_join(chaco_spi, by = "date") %>% 
  rename(chaco = spi) %>% 
  left_join(chinle_spi, by = "date") %>% 
  rename(chinle = spi) %>% 
  left_join(up_pue_spi, by = "date") %>% 
  rename(upper_puerco = spi) %>% 
  select(date, middle_sj, chaco, chinle, upper_puerco)
```

#### CHIRPS SPI vs Gridmet SPI
```{r, echo = FALSE}
# read in gridmet precip
chaco_precip_grid <- read_csv("watersheds/chaco_precip.csv") 
chinle_precip_grid <- read_csv("watersheds/chinle_precip.csv") 
mid_sj_precip_grid <- read_csv("watersheds/mid_sj_precip.csv") 
up_pue_precip_grid <- read_csv("watersheds/upper_puerco_precip.csv") 

#change to spi
mid_sj_spi_grid <- precip_to_spi(mid_sj_precip_grid)
chaco_spi_grid <- precip_to_spi(chaco_precip_grid)
chinle_spi_grid <- precip_to_spi(chinle_precip_grid)
up_pue_spi_grid <- precip_to_spi(up_pue_precip_grid)
```

```{r, echo = FALSE}
# make conglomerate of gridmet spi
watersheds_spi_grid <- mid_sj_spi_grid %>% 
  rename(middle_sj_grid = spi) %>% 
  left_join(chaco_spi_grid, by = "date") %>% 
  rename(chaco_grid = spi) %>% 
  left_join(chinle_spi_grid, by = "date") %>% 
  rename(chinle_grid = spi) %>% 
  left_join(up_pue_spi_grid, by = "date") %>% 
  rename(upper_puerco_grid = spi) %>% 
  select(date, middle_sj_grid, chaco_grid, chinle_grid, upper_puerco_grid) 

# combine it with the other watershed data
watersheds_spi_grid_chirps <- watersheds_spi_grid %>% 
  filter(year(date) >= 1982) %>% 
  left_join(watersheds_spi, by = "date") %>% 
  column_to_rownames("date") %>% 
  drop_na()
```

```{r, echo = FALSE}
# correlation matrix

# make correlation matrix for april 
cor_spi_chirps_grid <- cor(watersheds_spi_grid_chirps)
shirps_grid_p.mat <- cor.mtest(watersheds_spi_grid_chirps)$p # get p values

# makes correlations that hav an insignificant p value have a cross
corrplot(cor_spi_chirps_grid, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45,
         p.mat = shirps_grid_p.mat, sig.level = 0.05)

```


#### April SPI and Chuska Watershed Anomaly Correlations
```{r, echo = FALSE}
# April SPI combined with chuska winter anomaly
ws_ch_apr <- watersheds_spi %>% 
  filter(month(date) == 4) %>% 
  add_water_year() %>% 
  select(-date) %>% 
  filter(waterYear >= 2004) %>% 
  left_join(ch_wint_anom, by = "waterYear") %>% 
  rename(chuska_anom = anomaly_perc) %>% 
  column_to_rownames(var = "waterYear")
```

```{r, echo = FALSE}
# make correlation matrix for april 
cor_spi_ch_apr <- cor(ws_ch_apr)
ch_apr_p.mat <- cor.mtest(ws_ch_apr)$p # get p values

# makes correlations that hav an insignificant p value have a cross
corrplot(cor_spi_ch_apr, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45,
         p.mat = ch_apr_p.mat, sig.level = 0.05)
```

- Middle San Juan April SPI has the highest correlation with Chuska winter swe anom, chinle has the lowest
- Chinle also has the lowest correlations with other watersheds' SPIs

#### May SPI and Chuska Watershed Anomaly Correlations
```{r, echo = FALSE}
# April SPI combined with chuska winter anomaly
ws_ch_may <- watersheds_spi %>% 
  filter(month(date) == 5) %>% 
  add_water_year() %>% 
  select(-date) %>% 
  filter(waterYear >= 2004) %>% 
  left_join(ch_wint_anom, by = "waterYear") %>% 
  rename(chuska_anom = anomaly_perc) %>% 
  column_to_rownames(var = "waterYear")
```

```{r, echo = FALSE}
cor_spi_ch_may <- cor(ws_ch_may)
corrplot(cor_spi_ch_may, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)
```

- the correlations for may spi across all regions is much lower than for april
- Middle San Juan still had the higest correlation to Chuska, and Chinle has the least

### Middle San Juan

#### SPI
```{r, echo = FALSE}
# look at it
plot_spi(mid_sj_spi, month = all_months, title = "", years = "all") 

# the linear model info:
mid_sj_spi_lm <- lm(mid_sj_spi$spi ~ mid_sj_spi$date)
summary(mid_sj_spi_lm)
```

#### Chuska Monthly Winter SWE Anomaly vs SPI

```{r, echo = FALSE, message = FALSE}
# let's look at the correlations between SPI and chuska swe anom for April-July SPI 
plot_spi_swe_all(mid_sj_spi, ch_wint_anom)

```


- April SPI has the highest correlation (almost significantly!)

#### April SPI and Chuska swe
```{r, echo = FALSE}
# dataframe of april spi and chuska
mid_sj_spi_apr <- merge_ch_spi(mid_sj_spi, df_swe = ch_wint_anom, months = c(4)) 

# swe and spi following for April spi
plot_swe_spi(mid_sj_spi, month = c(4), swe_df = ch_wint_anom, region_time = "Navajo Nation April 6 month")

mid_sj_correct_apr <- swe_spi_same(mid_sj_spi_apr, type = "total")
mid_sj_positive_apr <- swe_spi_same(mid_sj_spi_apr, type = "positive")
mid_sj_negative_apr <- swe_spi_same(mid_sj_spi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(mid_sj_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r mid_sj_negative_apr*100`%
- *April* total correct: `r mid_sj_correct_apr*100`%

```{r, echo = FALSE}
plot_spi_drought_level(mid_sj_spi, 
                       month = c(4), 
                       region_time = "Navajo Nation April 6 month") 
```

- drought emergency in 2012 and 2018

```{r, echo = FALSE}
# chuska winter swe anom and april spi 
mid_sj_drought_times_all <- mid_sj_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  mutate(watershed = "middle_sj") %>% 
  select(waterYear, watershed, spi, swe_anom, drought)

mid_sj_drought_times <- mid_sj_drought_times_all %>% 
    filter(drought == "warning" | drought == "emergency") 
mid_sj_drought_times
```


```{r, echo = FALSE}

# graph of swe anomaly with the color relating to what level SPI drought warning follows it
mid_sj_swe_drought_apr <- mid_sj_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  ggplot(aes(x = waterYear, y = swe_anom,
             fill = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal"))
               )) +
  geom_col() +
  scale_fill_manual(values = c("#d13111", 
                                "#ffa600", 
                                "#f1d333", 
                                "#16990c"), 
                     labels = c("Drought Emergency", 
                                "Drought Warning", 
                                "Drought Alert", 
                                "Normal")) +
  labs(
    x = "Water Year",
    y = "SWE",
    fill = "Drought Severity",
    title = "Winter Swe Anomaly colored by April SPI drought warning"
  ) +
  theme_classic()

mid_sj_swe_drought_apr
```


### Chaco 


#### SPI
```{r, echo = FALSE}
# look at it
plot_spi(chaco_spi, month = all_months, title = "", years = "all") 

# the linear model info:
chaco_spi_lm <- lm(chaco_spi$spi ~ chaco_spi$date)
summary(chaco_spi_lm)
```

#### Chuska Monthly Winter SWE Anomaly vs SPI

```{r, echo = FALSE, message = FALSE}
# let's look at the correlations between SPI and chuska swe anom for April-July SPI 
plot_spi_swe_all(chaco_spi, ch_wint_anom)

```

- April SPI has the highest correlation (almost significantly!)

#### April SPI and Chuska swe
```{r, echo = FALSE}
# dataframe of april spi and chuska
chaco_spi_apr <- merge_ch_spi(chaco_spi, df_swe = ch_wint_anom, months = c(4)) 

# swe and spi following for April spi
plot_swe_spi(chaco_spi, month = c(4), swe_df = ch_wint_anom, region_time = "Navajo Nation April 6 month")

chaco_correct_apr <- swe_spi_same(chaco_spi_apr, type = "total")
chaco_positive_apr <- swe_spi_same(chaco_spi_apr, type = "positive")
chaco_negative_apr <- swe_spi_same(chaco_spi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(chaco_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r chaco_negative_apr*100`%
- *April* total correct: `r chaco_correct_apr*100`%

```{r, echo = FALSE}
plot_spi_drought_level(chaco_spi, 
                       month = c(4), 
                       region_time = "Navajo Nation April 6 month") 
```
- drought emergencies in 2006 and 2018

```{r, echo = FALSE}
# chuska winter swe anom and april spi 
chaco_drought_times_all <- chaco_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  mutate(watershed = "chaco") %>% 
  select(waterYear, watershed, spi, swe_anom, drought)

chaco_drought_times <- chaco_drought_times_all %>% 
  filter(drought == "warning" | drought == "emergency") 
chaco_drought_times

```

## Upper Puerco

#### Chuska Monthly Winter SWE Anomaly vs SPI

```{r, echo = FALSE, message = FALSE}
# let's look at the correlations between SPI and chuska swe anom for April-July SPI 
plot_spi_swe_all(up_pue_spi, ch_wint_anom)

```

#### April SPI and Chuska swe
```{r, echo = FALSE}
# dataframe of april spi and chuska
up_pue_spi_apr <- merge_ch_spi(up_pue_spi, df_swe = ch_wint_anom, months = c(4)) 

# swe and spi following for April spi
plot_swe_spi(up_pue_spi, month = c(4), swe_df = ch_wint_anom, region_time = "Navajo Nation April 6 month")

up_pue_correct_apr <- swe_spi_same(up_pue_spi_apr, type = "total")
up_pue_positive_apr <- swe_spi_same(up_pue_spi_apr, type = "positive")
up_pue_negative_apr <- swe_spi_same(up_pue_spi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(up_pue_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r up_pue_negative_apr*100`%
- *April* total correct: `r up_pue_correct_apr*100`%

### Drought severity in spring
```{r, echo = FALSE}
plot_spi_drought_level(up_pue_spi, 
                       month = c(4), 
                       region_time = "Navajo Nation March, April, May 6 month") 
```

- there was a drought emergency in 2006 & 2018


```{r, echo = FALSE}
# chuska winter swe anom and april spi 
up_pue_drought_times_all <- up_pue_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  mutate(watershed = "upper_puerco") %>% 
  select(waterYear, watershed, spi, swe_anom, drought)

up_pue_drought_times <- up_pue_drought_times_all %>% 
    filter(drought == "warning" | drought == "emergency") 
up_pue_drought_times


```


## Chinle

#### Precipitation
```{r, echo = FALSE}

ggplot(chinle_precip, aes(x = date, y = precipitation)) +
  geom_col()

```


#### SPI

```{r, echo = FALSE}
# look at it
plot_spi(chinle_spi, month = all_months, title = "", years = "all") 

# the linear model info:
chinle_spi_lm <- lm(chinle_spi$spi ~ chinle_spi$date)
summary(chinle_spi_lm)
```


#### Chuska Monthly Winter SWE Anomaly vs SPI

```{r, echo = FALSE, message = FALSE}
# let's look at the correlations between SPI and chuska swe anom for April-July SPI 
plot_spi_swe_all(chinle_spi, ch_wint_anom)

```

#### April SPI and Chuska swe
```{r, echo = FALSE}
# dataframe of april spi and chuska
chinle_spi_apr <- merge_ch_spi(chinle_spi, df_swe = ch_wint_anom, months = c(4)) 

# swe and spi following for April spi
plot_swe_spi(chinle_spi, month = c(4), swe_df = ch_wint_anom, region_time = "Navajo Nation April 6 month")

chinle_correct_apr <- swe_spi_same(chinle_spi_apr, type = "total")
chinle_positive_apr <- swe_spi_same(chinle_spi_apr, type = "positive")
chinle_negative_apr <- swe_spi_same(chinle_spi_apr, type = "negative")
```

- *April* positive swe & positive spi / positive swe: `r round(chinle_positive_apr*100, 3)`%
- *April* negative swe & negative spi / negative swe: `r chinle_negative_apr*100`%
- *April* total correct: `r chinle_correct_apr*100`%

### Drought severity in spring
```{r}
plot_spi_drought_level(chinle_spi, 
                       month = c(4), 
                       region_time = "Navajo Nation March, April, May 6 month") 

plot_spi_drought_level(chinle_spi, 
                       month = all_months, 
                       region_time = "Navajo Nation March, April, May 6 month",
                       years = "all")
```

- drought emergency occured in 2006, 2012, 2018

```{r, echo = FALSE}

chinle_drought_times_all <- chinle_spi_apr %>% 
  select(waterYear, spi, swe_anom) %>% 
  add_drought_severity() %>% 
  mutate(watershed = "chinle") %>% 
  select(waterYear, watershed, spi, swe_anom, drought)

chinle_drought_times <- chinle_drought_times_all %>% 
  filter(drought == "warning" | drought == "emergency")
chinle_drought_times

```



```{r, echo = FALSE}
## Make a dataframe of the percent right
watersheds = c("Middle San Juan", "Chaco", "Upper Puerco", "Chinle")

# ratio of times a negative swe anomaly was followed by a negative spi value #  out of the number of times there was a negative swe anomaly
negative_apr = c(mid_sj_negative_apr$n, chaco_negative_apr$n,
                 up_pue_negative_apr$n, chinle_negative_apr$n)

# ratio of times a positive swe anomaly was followed by a positive spi value #  out of the number of times there was a positive swe anomaly
positive_apr = c(mid_sj_positive_apr$n, chaco_positive_apr$n,
                 up_pue_positive_apr$n, chinle_positive_apr$n)

# ratio of the times the signs of swe and spi were the same
total_apr = c(mid_sj_correct_apr$n, chaco_correct_apr$n,
                 up_pue_correct_apr$n, chinle_correct_apr$n)

ws_same_sign = data.frame(watershed = c(watersheds),
                          april_negative = c(negative_apr),
                          april_positive = c(positive_apr),
                          april_total = c(total_apr))

ws_same_sign


```


### Relationships between CHuska Swe ANom and April watershed SPI distributions

```{r, echo = FALSE}
# bind the number of times 
drought_times <- rbind(up_pue_drought_times, chaco_drought_times,
                       mid_sj_drought_times, chinle_drought_times)

drought_times_all <- rbind(up_pue_drought_times_all, chaco_drought_times_all,
                       mid_sj_drought_times_all, chinle_drought_times_all)


# anova test

summary(aov(swe_anom ~ drought, data = drought_times_all))

```

```{r}
## let's test variances
# what are the variances
drought_variances <- drought_times_all %>% 
  group_by(drought) %>% 
  summarize(standard_deviation = round(sd(swe_anom),3),
            variance = round(var(swe_anom),3)) %>% 
  arrange(variance)

# LEvene's test to see if variances are signficiantly different
drought_levene <- leveneTest(swe_anom ~ drought, 
                             data = drought_times_all)
# the p value < 0.001 so variances are not equal 


```


```{r, echo = FALSE}
# visualize correlation 
drought_times_graph <- drought_times_all %>% 
  ggplot(aes(x = swe_anom, y = spi)) +
  geom_point(aes(color = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal")),
                 shape = watershed)) +
  scale_color_manual(values = c("#d13111", 
                                  "#ffa600", 
                                  "#f1d333", 
                                  "#16990c")) +
  labs(
    x = "Chuska SWE Winter Anomaly",
    y = "April SPI",
    color = "Drought Severity"
  ) +
  theme_classic() 
drought_times_graph
```


```{r, echo = FALSE}
drought_times_boxplot <- drought_times_all %>% 
  ggplot(aes(x = watershed, y = swe_anom)) +
  geom_boxplot(aes(fill = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal")
                             )
                   )
               )+
  scale_fill_manual(values = c("#d13111", 
                                  "#ffa600", 
                                  "#f1d333", 
                                  "#16990c")) +
  labs(
    x = "Watershed",
    y = "Chuska Winter Swe Anomaly",
    fill = "Drought"
  ) +
  theme_classic()

drought_times_boxplot
```

```{r, echo = FALSE}

drought_times_boxplot_all <- drought_times_all %>% 
  ggplot(aes(x = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal")), 
                        y = swe_anom)) +
  geom_boxplot(aes(color = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal"))
                   ))+
  geom_jitter(aes(color = factor(drought, 
                             levels = c("emergency",
                                        "warning",
                                        "alert",
                                        "normal"))
                   )) +
  scale_color_manual(values = c("#d13111", 
                                  "#ffa600", 
                                  "#f1d333", 
                                  "#16990c")) +
  labs(
    x = "Watershed",
    y = "Chuska Winter Swe Anomaly",
    color = "Drought"
  ) +
  theme_classic()

drought_times_boxplot_all

```

```{r, echo = FALSE}
# do some stats to see how often different things occur

drought_swe_levels <- drought_times_all %>% 
  filter(swe_anom >= 0.9) %>% 
  count(drought) %>% 
  mutate(perc = n/sum(n))

```

### Tsaile/wheatfields

```{r, echo = FALSE}
tsaile_spi <- precip_to_spi(tsaile_precip)
```

```{r, echo = FALSE}
# look at the SPI
plot_spi_drought_level(tsaile_spi, 
                       month = all_months, 
                       region_time = "Tsaile",
                       years = "all")
```

```{r, echo = FALSE}
# compare to chuska swe
plot_spi_swe_all(tsaile_spi, ch_wint_anom)

```

```{r, echo = FALSE}

tsaile_spi_apr <- merge_ch_spi(tsaile_spi, 
                               df_swe = ch_wint_anom, 
                               months = c(4)) 

tsaile_correct_apr <- swe_spi_same(tsaile_spi_apr, type = "total")
tsaile_positive_apr <- swe_spi_same(tsaile_spi_apr, type = "positive")
tsaile_negative_apr <- swe_spi_same(tsaile_spi_apr, type = "negative")
```


### Day of year of snowmelt and SPI
```{r, echo = FALSE}
# average day of year of chuska snowmelt
ch_snowmelt_graph <- ch_snowmelt %>% 
  rename(mean_doy = doy_snowmelt) %>% 
  melt(id.vars = "year") %>% 
  rename(snowmelt_doy = value) %>% 
  ggplot(aes(x = year, y = snowmelt_doy)) +
  geom_line(aes(color = variable)) +
  theme_classic()
ch_snowmelt_graph
```

#### April
```{r, echo = FALSE}
# look at SPI watersheds and day of snowmelt
spi_snowmelt_apr <- watersheds_spi %>% 
  filter(year(date) >= 2001) %>% 
  filter(year(date) <= 2015) %>% 
  filter(month(date) == 4) %>% 
  mutate(year = year(date)) %>% 
  select(year, middle_sj, chaco, chinle, upper_puerco) %>% 
  left_join(ch_snowmelt, by = "year") %>% 
  column_to_rownames("year")

# correlation matrix
cor_spi_snowmelt_apr <- cor(spi_snowmelt_apr)
corrplot(cor_spi_snowmelt_apr, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)
```
#### May
```{r, echo = FALSE}
# look at SPI watersheds and day of snowmelt
spi_snowmelt_may <- watersheds_spi %>% 
  filter(year(date) >= 2001) %>% 
  filter(year(date) <= 2015) %>% 
  filter(month(date) == 5) %>% 
  mutate(year = year(date)) %>% 
  select(year, middle_sj, chaco, chinle, upper_puerco) %>% 
  left_join(ch_snowmelt, by = "year") %>% 
  column_to_rownames("year")

# correlation matrix
cor_spi_snowmelt_may <- cor(spi_snowmelt_may)
p.mat <- cor.mtest(spi_snowmelt_may)$p
corrplot(cor_spi_snowmelt_may, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45,
         p.mat = p.mat, sig.level = 0.05)
```

```{r, echo = FALSE}
# look at SPI watersheds and day of snowmelt
spi_snowmelt_june <- watersheds_spi %>% 
  filter(year(date) >= 2001) %>% 
  filter(year(date) <= 2015) %>% 
  filter(month(date) == 6) %>% 
  mutate(year = year(date)) %>% 
  select(year, middle_sj, chaco, chinle, upper_puerco) %>% 
  left_join(ch_snowmelt, by = "year") %>% 
  column_to_rownames("year")

# correlation matrix
cor_spi_snowmelt_june <- cor(spi_snowmelt_june)
corrplot(cor_spi_snowmelt_june, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45)
```


```{r, echo = FALSE}
# Chuska SWE average swe and doy snowmelt
swe_snowmelt <- ch_wint_anom %>% 
  filter(waterYear <= 2015) %>% 
  rename(year = waterYear) %>% 
  left_join(ch_snowmelt, by = "year") %>% 
  column_to_rownames("year") %>% 
  rename(mean_doy = doy_snowmelt,
         swe_anomaly = anomaly_perc)
  
swe_snowmelt_cor <- cor(swe_snowmelt)
swe_snowmelt_p.mat <- cor.mtest(swe_snowmelt)$p
corrplot(swe_snowmelt_cor, method = "ellipse", type = "upper", 
         addCoef.col = "black", tl.col = "black",  tl.srt=45,
         p.mat = swe_snowmelt_p.mat, sig.level = 0.05)
```

```{r, echo = FALSE}
ggplot(swe_snowmelt, aes(x = swe_anomaly, y = mean_doy)) +
  geom_point() +
  theme_classic()

```

