---
title: "Standardized Precipitation Index"
author: "AnnaClaire Marley"
date: "7/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
library(precintcon)
source("Functions/util.R")
source('Functions/graphing.R')
```

This Rmarkdown shows analysis of SPI at the watersheds bordering the Chuskas and Tsaile chaper

```{r, echo = FALSE, message=FALSE}
# first load in chuska winter anomalies

ch_wint_anom <- read_csv("Chuska/ch_wint_swe_anom.csv") %>% 
  select(waterYear, anomaly_perc)

# for later use:
all_months = c(1,2,3,4,5,6,7,8,9,10,11,12)

```

# Watersheds

```{r, include = FALSE}
# read in all the watersheds precipitation data
chaco_precip <- read_csv("watersheds/chaco_precip.csv") 
chinle_precip <- read_csv("watersheds/chinle_precip.csv") 
mid_sj_precip <- read_csv("watersheds/mid_sj_precip.csv") 
up_pue_precip <- read_csv("watersheds/upper_puerco_precip.csv") 

```


```{r, echo = FALSE}
# read in the watersheds' monthly pdsi data to compare to spi
chaco_pdsi <- read_csv("watersheds/chaco_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year() %>% 
  calc_pdsi_metrics("month", mean) 

chinle_pdsi <- read_csv("watersheds/chinle_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year() %>% 
  calc_pdsi_metrics("month", mean)

up_pue_pdsi <- read_csv("watersheds/upper_puerco_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year() %>% 
  calc_pdsi_metrics("month", mean)

mid_sj_pdsi <- read_csv("watersheds/middle_san_juan_pdsi.csv", col_names=c("date", "pdsi"), skip =1) %>% 
  add_water_year() %>% 
  calc_pdsi_metrics("month", mean)

```

## Chaco

### Chaco SPI

```{r, echo = FALSE, message=FALSE}

# now change the precip data to spi
chaco_spi <- precip_to_spi(chaco_precip)


```

```{r, echo = FALSE}

# whole year 6 month SPI
plot_spi(chaco_spi, month = all_months, title = "Chaco SPI", years = "all") 



# let's take a look at the 6 month June SPI of each year
plot_spi(chaco_spi, month = c(6), title = "Chaco June SPI")

```

### Chaco PDSI and SPI

```{r, echo = FALSE}
chaco_pdsi_spi <- chaco_spi %>% 
  select(date, spi) %>% 
  merge(chaco_pdsi, by = "date") %>% 
  rename(mean_pdsi = pdsi)

plot_cor_pdsi_spi(chaco_pdsi_spi, title = "Chaco")

```


m### Chaco SPI and Chuska SWE

```{r, echo = FALSE}
### now let's look at Chuska monthly swe anomaly and 6 month SPI ###

plot_swe_spi(chaco_spi, month = c(6), region_time = "Chaco June 6 month")
```

```{r, echo = FALSE}

# let's look at the correlation between them 
ch_chaco_spi_june <- merge_ch_spi(chaco_spi, months = c(6)) 
ch_chaco_spi_june_plot <- plot_cor_swe_spi(ch_chaco_spi_june,
                                           title = "Winter SWE Anomaly & June SPI")


ch_chaco_correct <- swe_spi_same(ch_chaco_spi_june, type = "total")
ch_chaco_positive <- swe_spi_same(ch_chaco_spi_june, type = "positive")
ch_chaco_negative <- swe_spi_same(ch_chaco_spi_june, type = "negative")

```

- The percent of times swe the anomaly for the year and spi had the same sign `r ch_chaco_correct*100`%

- The percent of times swe anomaly for the year was positive and was followed by a positive spi had the same sign out of all positive swe years `r ch_chaco_positive*100`%

- The percent of times swe anomaly for the year was negative and was followed by a negative spi had the same sign out of all negative swe years `r ch_chaco_negative*100`%

```{r, echo = FALSE}
# let's try for july spi

ch_chaco_spi_july <- merge_ch_spi(chaco_spi, months = c(7)) 

ch_chaco_spi_july_plot <- plot_cor_swe_spi(ch_chaco_spi_july, 
                                           title = "Winter SWE Anomaly & July SPI")


```



```{r, echo = FALSE}
# let's try for May spi

ch_chaco_spi_may <- merge_ch_spi(chaco_spi, months = c(5))
ch_chaco_spi_may_plot <- plot_cor_swe_spi(ch_chaco_spi_may, title = "Winter SWE Anomaly & May SPI")


```

```{r, echo = FALSE}

# let's look at the correlation between them  for MAY
ch_chaco_correct_may <- swe_spi_same(ch_chaco_spi_may, type = "total")
ch_chaco_positive_may <- swe_spi_same(ch_chaco_spi_may, type = "positive")
ch_chaco_negative_may <- swe_spi_same(ch_chaco_spi_may, type = "negative")

```


```{r, echo = FALSE}
# let's try for April spi

ch_chaco_spi_apr <- merge_ch_spi(chaco_spi, months = c(4)) 
  
ch_chaco_spi_apr_plot <- plot_cor_swe_spi(ch_chaco_spi_apr,
                                          title = "Winter SWE Anomaly & April SPI")

```

```{r, echo=FALSE, fig.width=7}

grid.arrange(ch_chaco_spi_apr_plot, ch_chaco_spi_may_plot, 
             ch_chaco_spi_june_plot, ch_chaco_spi_july_plot, ncol = 2)

```

- June SPI seems to be much more correlated to winter SWE than does July SPI
- May is the most well correlated to winter swe

```{r, echo = FALSE, error = TRUE}
plot_swe_spi(chaco_spi, month = c(5), region_time = "Chaco May 6 month")

```


- *May* positive swe & positive spi / positive swe: `r round(ch_chaco_positive_may*100, 3)`%
- *May* negative swe & negative spi / negative swe: `r ch_chaco_negative_may*100`%
- *May* total correct: `r ch_chaco_correct_may*100`%

## Chinle

### Chinle SPI

```{r, echo = FALSE, message=FALSE}
# read in precipita
# now change the precip data to spi
chinle_spi <- precip_to_spi(chinle_precip)

# SPI June

```

```{r, echo = FALSE}

# whole year 6 month SPI
plot_spi(chinle_spi, month = all_months, title = "Chinle SPI", years = "all")

# let's take a look at the 6 month June SPI of each year
plot_spi(chinle_spi, month = c(6), title = "Chinle June SPI")

```

### Chinle SPI vs PDSI

```{r, echo = FALSE}
chinle_pdsi_spi <- chinle_spi %>% 
  select(date, spi) %>% 
  merge(chinle_pdsi, by = "date") %>% 
  rename(mean_pdsi = pdsi)

plot_cor_pdsi_spi(chinle_pdsi_spi, title = "Chinle")

```

### Chinle SPI and Chuska SWE

```{r, echo = FALSE}
### now let's look at Chuska monthly swe anomaly and June 6 month SPI ###
plot_swe_spi(chinle_spi, month = c(6), region_time = "Chinle June 6 month")

```

```{r, echo = FALSE}
# let's look at the correlation between them 
ch_chinle_spi_june <- merge_ch_spi(chinle_spi, months = c(6))
ch_chinle_spi_june_plot <- plot_cor_swe_spi(ch_chinle_spi_june, 
                                            title = "Winter SWE Anomaly & June SPI")

ch_chinle_correct <- swe_spi_same(ch_chinle_spi_june, type = "total")
ch_chinle_positive <- swe_spi_same(ch_chinle_spi_june, type = "positive")
ch_chinle_negative <- swe_spi_same(ch_chinle_spi_june, type = "negative")

```

- The percent of times swe the anomaly for the year and spi had the same sign `r ch_chinle_correct*100`%

- The percent of times swe anomaly for the year was positive and was followed by a positive spi had the same sign out of all positive swe years `r ch_chinle_positive*100`%

- The percent of times swe anomaly for the year was negative and was followed by a negative spi had the same sign out of all negative swe years `r ch_chinle_negative*100`%

```{r, echo=FALSE}
# try July SPI
ch_chinle_spi_july <- merge_ch_spi(chinle_spi, months = c(7)) 
ch_chinle_spi_july_plot <- plot_cor_swe_spi(ch_chinle_spi_july, 
                                      title = "Winter SWE Anomaly & July SPI")


```


```{r, echo=FALSE}
# try May SPI
ch_chinle_spi_may <- merge_ch_spi(chinle_spi, months = c(5)) 
ch_chinle_spi_may_plot <- plot_cor_swe_spi(ch_chinle_spi_may, 
                                           title = "Winter SWE Anomaly & May SPI")

# April:
ch_chinle_spi_apr <- merge_ch_spi(chinle_spi, months = c(4)) 
ch_chinle_spi_apr_plot <- plot_cor_swe_spi(ch_chinle_spi_apr, 
                                           title = "Winter SWE Anomaly & April SPI")

```

```{r, echo = FALSE, fig.width=7}

grid.arrange(ch_chinle_spi_apr_plot, ch_chinle_spi_may_plot, 
             ch_chinle_spi_june_plot, ch_chinle_spi_july_plot, ncol = 2)
```

- Chuska SWE is much less corelated to Chinle July 6 month SPI than it is to June 6 month
- May had the higest correlation

```{r, echo = FALSE}
plot_swe_spi(chinle_spi, month = c(5), region_time = "Chinle May 6 month")

ch_chinle_correct_may <- swe_spi_same(ch_chinle_spi_may, type = "total")
ch_chinle_positive_may <- swe_spi_same(ch_chinle_spi_may, type = "positive")
ch_chinle_negative_may <- swe_spi_same(ch_chinle_spi_may, type = "negative")

```

- *May* positive swe & positive spi / positive swe: `r round(ch_chinle_positive_may*100, 3)`%
- *May* negative swe & negative spi / negative swe: `r ch_chinle_negative_may*100`%
- *May* total correct: `r ch_chinle_correct_may*100`%

## Middle San Juan

```{r, echo = FALSE, message=FALSE}
# read in precipitation


# now change the precip data to spi
mid_sj_spi <- precip_to_spi(mid_sj_precip)


```

```{r, echo = FALSE}
# whole year 6 month SPI
plot_spi(mid_sj_spi, month = all_months, title = "Middle SJ SPI", years = "all")

# let's take a look at the 6 month June SPI of each year
plot_spi(mid_sj_spi, month = c(6), title = "Middle SJ June SPI")

```

### Middle San Juan SPI and PDSI

```{r, echo = FALSE}
mid_sj_pdsi_spi <- mid_sj_spi %>% 
  select(date, spi) %>% 
  merge(mid_sj_pdsi, by = "date") %>% 
  rename(mean_pdsi = pdsi)

plot_cor_pdsi_spi(mid_sj_pdsi_spi, title = "Middle San Juan")

```
### Middle San Juan SPI and Chuska SWE

```{r, echo = FALSE}
### now let's look at Chuska monthly swe anomaly and June 6 month SPI ###
plot_swe_spi(mid_sj_spi, month = c(6), region_time = "Middle San Juan June 6 month")

```

```{r, echo = FALSE}
# let's look at the correlation between them 
ch_mid_sj_spi_june <- merge_ch_spi(mid_sj_spi, months = c(6))

ch_mid_sj_spi_june_plot <- plot_cor_swe_spi(ch_mid_sj_spi_june, 
                 title = "Winter SWE Anomaly & June SPI")

ch_mid_sj_correct <- swe_spi_same(ch_mid_sj_spi_june, type = "total")
ch_mid_sj_positive <- swe_spi_same(ch_mid_sj_spi_june, type = "positive")
ch_mid_sj_negative <- swe_spi_same(ch_mid_sj_spi_june, type = "negative")

```

- The percent of times swe the anomaly for the year and spi had the same sign `r ch_mid_sj_correct*100`%

- The percent of times swe anomaly for the year was positive and was followed by a positive spi had the same sign out of all positive swe years `r ch_mid_sj_positive*100`%

- The percent of times swe anomaly for the year was negative and was followed by a negative spi had the same sign out of all negative swe years `r ch_mid_sj_negative*100`%

```{r, echo = FALSE}
ch_mid_sj_spi_apr <- merge_ch_spi(mid_sj_spi, months = c(4))

ch_mid_sj_spi_apr_plot <- plot_cor_swe_spi(ch_mid_sj_spi_apr, 
                 title = "Winter SWE Anomaly & April SPI")

ch_mid_sj_spi_may <- merge_ch_spi(mid_sj_spi, months = c(5))

ch_mid_sj_spi_may_plot <- plot_cor_swe_spi(ch_mid_sj_spi_may, 
                 title = "Winter SWE Anomaly & May SPI")

ch_mid_sj_spi_july <- merge_ch_spi(mid_sj_spi, months = c(7))

ch_mid_sj_spi_july_plot <- plot_cor_swe_spi(ch_mid_sj_spi_july, 
                 title = "Winter SWE Anomaly & July SPI")

grid.arrange(ch_mid_sj_spi_apr_plot, ch_mid_sj_spi_may_plot, 
             ch_mid_sj_spi_june_plot, ch_mid_sj_spi_july_plot, ncol=2)
```

```{r, echo = FALSE}
# swe and spi following for MAY spi
plot_swe_spi(mid_sj_spi, month = c(5), region_time = "Middle San Juan May 6 month")

ch_mid_sj_correct_may <- swe_spi_same(ch_mid_sj_spi_may, type = "total")
ch_mid_sj_positive_may <- swe_spi_same(ch_mid_sj_spi_may, type = "positive")
ch_mid_sj_negative_may <- swe_spi_same(ch_mid_sj_spi_may, type = "negative")
```


- *May* positive swe & positive spi / positive swe: `r round(ch_mid_sj_positive_may*100, 3)`%
- *May* negative swe & negative spi / negative swe: `r ch_mid_sj_negative_may*100`%
- *May* total correct: `r ch_mid_sj_correct_may*100`%

## Upper  Puerco

```{r, echo = FALSE, message=FALSE}

# now change the precip data to spi
up_pue_spi <- precip_to_spi(up_pue_precip)

```

```{r, echo = FALSE}
# whole year 6 month SPI
plot_spi(up_pue_spi, month = all_months, title = "Upper Puerco SPI", years = "all")

# let's take a look at the 6 month June SPI of each year
plot_spi(up_pue_spi, month = c(6), title = "Upper Puerco June SPI")

```

### Upper Puerco San Juan SPI and PDSI


```{r, echo = FALSE}
up_pue_pdsi_spi <- up_pue_spi %>% 
  select(date, spi) %>% 
  merge(up_pue_pdsi, by = "date") %>% 
  rename(mean_pdsi = pdsi)

plot_cor_pdsi_spi(up_pue_pdsi_spi, title = "Upper Puerco")

```

### Upper Puerco SPI and Chuska SWE

```{r, echo = FALSE}
ch_up_pue_spi_apr <- merge_ch_spi(up_pue_spi, months = c(4)) 

ch_up_pue_spi_apr_plot <- plot_cor_swe_spi(ch_up_pue_spi_apr, 
                 title = "Winter SWE Anomaly & April SPI")

ch_up_pue_spi_may <- merge_ch_spi(up_pue_spi, months = c(5)) 

ch_up_pue_spi_may_plot <- plot_cor_swe_spi(ch_up_pue_spi_may, 
                 title = "Winter SWE Anomaly & May SPI")

ch_up_pue_spi_june <- merge_ch_spi(up_pue_spi, months = c(6))

ch_up_pue_spi_june_plot <- plot_cor_swe_spi(ch_up_pue_spi_june, 
                 title = "Winter SWE Anomaly & June SPI")

ch_up_pue_spi_july <- merge_ch_spi(up_pue_spi, months = c(7))

ch_up_pue_spi_july_plot <- plot_cor_swe_spi(ch_up_pue_spi_july, 
                 title = "Winter SWE Anomaly & July SPI")

```

```{r, echo = FALSE}

grid.arrange(ch_up_pue_spi_apr_plot, ch_up_pue_spi_may_plot, 
             ch_up_pue_spi_june_plot, ch_up_pue_spi_july_plot, ncol=2)
```

- May has the highest correlation

```{r, echo = FALSE}
# without the 2010 water year
ch_up_pue_spi_apr_n10 <- merge_ch_spi(up_pue_spi, months = c(4)) %>% 
  filter(waterYear != 2010)

ch_up_pue_spi_apr_plot_n10 <- plot_cor_swe_spi(ch_up_pue_spi_apr_n10, 
                 title = "Winter SWE Anomaly & April SPI")

ch_up_pue_spi_may_n10 <- merge_ch_spi(up_pue_spi, months = c(5)) %>% 
  filter(waterYear != 2010)

ch_up_pue_spi_may_plot_n10 <- plot_cor_swe_spi(ch_up_pue_spi_may_n10, 
                 title = "Winter SWE Anomaly & May SPI")

ch_up_pue_spi_june_n10 <- merge_ch_spi(up_pue_spi, months = c(6))%>% 
  filter(waterYear != 2010)

ch_up_pue_spi_june_plot_n10 <- plot_cor_swe_spi(ch_up_pue_spi_june_n10, 
                 title = "Winter SWE Anomaly & June SPI")

ch_up_pue_spi_july_n10 <- merge_ch_spi(up_pue_spi, months = c(7))%>% 
  filter(waterYear != 2010)

ch_up_pue_spi_july_plot_n10 <- plot_cor_swe_spi(ch_up_pue_spi_july_n10, 
                 title = "Winter SWE Anomaly & July SPI")

```

#### Removing 2010 from the correlation
```{r, echo = FALSE}

grid.arrange(ch_up_pue_spi_apr_plot_n10, ch_up_pue_spi_may_plot_n10, 
             ch_up_pue_spi_june_plot_n10, ch_up_pue_spi_july_plot_n10, ncol=2)
```

- taking out the 2010 water year does improve the correlation, May r -> 0.436, April r -> 0.385

```{r, echo = FALSE}
# swe and spi following for MAY spi
plot_swe_spi(up_pue_spi, month = c(5), region_time = "Upper Puerco May 6 month")

ch_up_pue_correct_may <- swe_spi_same(ch_up_pue_spi_may, type = "total")
ch_up_pue_positive_may <- swe_spi_same(ch_up_pue_spi_may, type = "positive")
ch_up_pue_negative_may <- swe_spi_same(ch_up_pue_spi_may, type = "negative")
```


- *May* positive swe & positive spi / positive swe: `r round(ch_up_pue_positive_may*100, 3)`%
- *May* negative swe & negative spi / negative swe: `r ch_up_pue_negative_may*100`%
- *May* total correct: `r ch_up_pue_correct_may*100`%



----------------

## Tsaile Chapter

### Tsaile SPI

#### 6 month SPI
```{r, include = FALSE}
# read in TSAILE precipitation

tsaile_precip <- read_csv("Chapters/tsaile_precip.csv")

# now change the precip data to spi
tsaile_spi <- precip_to_spi(tsaile_precip)


```

```{r, echo = FALSE}

# whole year 6 month SPI
plot_spi(tsaile_spi, month = all_months, title = "Tsaile SPI", years = "all")

# let's take a look at the 6 month June SPI of each year
plot_spi(tsaile_spi, month = c(6), title = "Tsaile June SPI")

```

#### Tsaile SPI and Chuska SWE anomaly

```{r, echo = FALSE}
# plot correlations between swe and different month SPIs

ch_tsaile_spi_apr <- merge_ch_spi(tsaile_spi, months = c(4))

ch_tsaile_spi_apr_plot <- plot_cor_swe_spi(ch_tsaile_spi_apr, 
                 title = "Winter SWE Anomaly & April SPI")

ch_tsaile_spi_may <- merge_ch_spi(tsaile_spi, months = c(5))

ch_tsaile_spi_may_plot <- plot_cor_swe_spi(ch_tsaile_spi_may, 
                 title = "Winter SWE Anomaly & May SPI")

ch_tsaile_spi_june <- merge_ch_spi(tsaile_spi, months = c(6))

ch_tsaile_spi_june_plot <- plot_cor_swe_spi(ch_tsaile_spi_june, 
                 title = "Winter SWE Anomaly & June SPI")

ch_tsaile_spi_july <- merge_ch_spi(tsaile_spi, months = c(7))

ch_tsaile_spi_july_plot <- plot_cor_swe_spi(ch_tsaile_spi_july, 
                 title = "Winter SWE Anomaly & July SPI")

grid.arrange(ch_tsaile_spi_apr_plot, ch_tsaile_spi_may_plot, 
             ch_tsaile_spi_june_plot, ch_tsaile_spi_july_plot, ncol=2)

```
- May is has the higest correlation

### without 2010
```{r, echo = FALSE}
# plot correlations between swe and different month SPIs

ch_tsaile_spi_apr_n10 <- merge_ch_spi(tsaile_spi, months = c(4)) %>% 
  filter(waterYear != 2010)

ch_tsaile_spi_apr_plot_n10 <- plot_cor_swe_spi(ch_tsaile_spi_apr_n10, 
                 title = "Winter SWE Anomaly & April SPI")

ch_tsaile_spi_may_n10 <- merge_ch_spi(tsaile_spi, months = c(5)) %>% 
  filter(waterYear != 2010)

ch_tsaile_spi_may_plot_n10 <- plot_cor_swe_spi(ch_tsaile_spi_may_n10, 
                 title = "Winter SWE Anomaly & May SPI")

ch_tsaile_spi_june_n10 <- merge_ch_spi(tsaile_spi, months = c(6)) %>% 
  filter(waterYear != 2010)

ch_tsaile_spi_june_plot_n10 <- plot_cor_swe_spi(ch_tsaile_spi_june_n10, 
                 title = "Winter SWE Anomaly & June SPI")

ch_tsaile_spi_july_n10 <- merge_ch_spi(tsaile_spi, months = c(7)) %>% 
  filter(waterYear != 2010)

ch_tsaile_spi_july_plot_n10 <- plot_cor_swe_spi(ch_tsaile_spi_july_n10, 
                 title = "Winter SWE Anomaly & July SPI")

grid.arrange(ch_tsaile_spi_apr_plot_n10, ch_tsaile_spi_may_plot_n10, 
             ch_tsaile_spi_june_plot_n10, ch_tsaile_spi_july_plot_n10, ncol=2)

```

- taking out 2010 improves the correlation quite a bit

```{r, echo = FALSE}
# swe and spi following for MAY spi
plot_swe_spi(tsaile_spi, month = c(5), region_time = "Tsaile May 6 month")

ch_tsaile_spi_correct_may <- swe_spi_same(ch_tsaile_spi_may, type = "total")
ch_tsaile_spi_positive_may <- swe_spi_same(ch_tsaile_spi_may, type = "positive")
ch_tsaile_spi_negative_may <- swe_spi_same(ch_tsaile_spi_may, type = "negative")
```

- *May* positive swe & positive spi / positive swe: `r round(ch_tsaile_spi_positive_may*100, 3)`%
- *May* negative swe & negative spi / negative swe: `r ch_tsaile_spi_negative_may*100`%
- *May* total correct: `r ch_tsaile_spi_correct_may*100`%


```{r, echo = FALSE}
### now let's look at Chuska monthly swe anomaly and Tsaile SPI Jan-June ###
plot_swe_spi(tsaile_spi, month = c(6), region_time = "Tsaile June 6 month")

```

```{r, echo = FALSE}
# let's look at the correlation between them 

ch_tsaile_correct <- swe_spi_same(ch_tsaile_spi_june, type = "total")
ch_tsaile_positive <- swe_spi_same(ch_tsaile_spi_june, type = "positive")
ch_tsaile_negative <- swe_spi_same(ch_tsaile_spi_june, type = "negative")

```

- The percent of times swe the anomaly for the year and spi had the same sign `r ch_tsaile_correct*100`%

- The percent of times swe anomaly for the year was positive and was followed by a positive spi had the same sign out of all positive swe years `r ch_tsaile_positive*100`%

- The percent of times swe anomaly for the year was negative and was followed by a negative spi had the same sign out of all negative swe years `r ch_tsaile_negative*100`%

