---
title: "Standardized Precipitation Index"
author: "AnnaClaire Marley"
date: "7/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
library(corrplot)
library(precintcon)
source("Functions/util.R")
source('Functions/graphing.R')
```

This Rmarkdown shows analysis of SPI at the watersheds bordering the Chuskas and Tsaile chaper

```{r, echo = FALSE}
# first load in chuska winter anomalies

ch_wint_anom <- read_csv("Chuska/ch_wint_swe_anom.csv") %>% 
  select(waterYear, anomaly_perc)

# for later use:
all_months = c(1,2,3,4,5,6,7,8,9,10,11,12)

```

# Watersheds

## Chaco

### Chaco SPI

```{r, echo = FALSE}
# read in chaco precipitation

chaco_precip <- read_csv("watersheds/chaco_precip.csv") %>% 
  rename(Precipitation = `Precipitation (mm)`,
         date = DateTime)

# now change the precip data to spi
chaco_spi <- precip_to_spi(chaco_precip)

# Chaco SPI June

chaco_spi_june <- chaco_spi %>% 
  filter(year(date) >= 2004) %>% 
  filter(month(date) == 6) %>% 
  add_water_year() %>% 
  select(waterYear, spi)

```

```{r, echo = FALSE}

# whole year 6 month SPI
plot_spi(chaco_spi, month = all_months, title = "Chaco SPI")

# let's take a look at the 6 month June SPI of each year
plot_spi(chaco_spi, month = c(6), title = "Chaco June SPI")

```


### Chaco SPI and Chuska SWE

```{r, echo = FALSE}
### now let's look at Chuska monthly swe anomaly and June 6 month SPI ###
plot_swe_spi(chaco_spi, month = c(6), region_time = "Chaco June 6 month")

```


```{r, echo = FALSE}

# let's look at the correlation between them 
ch_chaco_spi_june <- chaco_spi_june %>% 
   merge(ch_wint_anom, by = "waterYear") %>% 
   rename(swe_anom = anomaly_perc)

plot_cor_swe_spi(ch_chaco_spi_june, title = "Chuska Winter SWE Anomaly and Tsaile Jan-June SPI")

ch_tsaile_correct <- swe_spi_same(ch_chaco_spi_june, type = "total")
ch_tsaile_positive <- swe_spi_same(ch_chaco_spi_june, type = "positive")
ch_tsaile_negative <- swe_spi_same(ch_chaco_spi_june, type = "negative")

```

- The percent of times swe the anomaly for the year and spi had the same sign `r ch_tsaile_correct*100`%

- The percent of times swe anomaly for the year was positive and was followed by a positive spi had the same sign out of all positive swe years `r ch_tsaile_positive*100`%

- The percent of times swe anomaly for the year was negative and was followed by a negative spi had the same sign out of all negative swe years `r ch_tsaile_negative*100`%

## Chinle

## Middle San Juan

## Upper  Puerco




## Tsaile Chapter

### Tsaile SPI

#### 6 month SPI
```{r, include = FALSE}
# read in TSAILE precipitation

tsaile_precip <- read_csv("Chapters/tsaile_precip.csv")

# now change the precip data to spi
tsaile_spi <- precip_to_spi(tsaile_precip)

# Tsaile SPI June

tsaile_spi_june <- tsaile_spi %>% 
  filter(year(date) >= 2004) %>% 
  filter(month(date) == 6) %>% 
  add_water_year() %>% 
  select(waterYear, spi)

```

```{r, echo = FALSE}

# whole year 6 month SPI
plot_spi(tsaile_spi, month = all_months, title = "Tsaile SPI")

# let's take a look at the 6 month June SPI of each year
plot_spi(tsaile_spi, month = c(6), title = "Tsaile June SPI")

```

#### 3 month SPI
```{r, include = FALSE}

# now change the precip data to spi
tsaile_spi_3 <- precip_to_spi(tsaile_precip, timeframe = 3)

```

```{r, echo = FALSE}


# whole year 6 month SPI
plot_spi(tsaile_spi, month = all_months, title = "Tsaile SPI")

# let's take a look at the 6 month June SPI of each year
plot_spi(tsaile_spi, month = c(6), title = "Tsaile June SPI")

```


### Tsaile SPI and Chuska SWE

```{r, echo = FALSE}
### now let's look at Chuska monthly swe anomaly and Tsaile SPI Jan-June ###
plot_swe_spi(tsaile_spi, month = c(6), region_time = "Tsaile June 6 month")

```

- 7/10 of negative swe years followed by negative spi
- 4/6 of positive swe years followed by positive spi 

```{r, echo = FALSE}

# let's look at the correlation between them 
ch_tsaile_spi_june <- tsaile_spi_june %>% 
   merge(ch_wint_anom, by = "waterYear") %>% 
   rename(swe_anom = anomaly_perc)

plot_cor_swe_spi(ch_tsaile_spi_june, title = "Chuska Winter SWE Anomaly and Tsaile Jan-June SPI")

ch_tsaile_correct <- swe_spi_same(ch_tsaile_spi_june, type = "total")
ch_tsaile_positive <- swe_spi_same(ch_tsaile_spi_june, type = "positive")
ch_tsaile_negative <- swe_spi_same(ch_tsaile_spi_june, type = "negative")

```

- The percent of times swe the anomaly for the year and spi had the same sign `r ch_tsaile_correct*100`%

- The percent of times swe anomaly for the year was positive and was followed by a positive spi had the same sign out of all positive swe years `r ch_tsaile_positive*100`%

- The percent of times swe anomaly for the year was negative and was followed by a negative spi had the same sign out of all negative swe years `r ch_tsaile_negative*100`%

### 3 month SPI
```{r, echo=FALSE}

plot_swe_spi(tsaile_spi_3, month = c(6), region_time = "Tsaile June 3 month")
```

