---
title: "SWE Final Analyses"
output: html_document
Author: AnnaClaire Marley
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown is a conglomerate of all the graphs created in SWE analyses from SNODAS data collected from Climate Engine

```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)

```

### Prelimary Analyses from data below:

- Chuska recieves the most snow out of the three
- 2010 is a big anomaly at all three locations, was this a heavy precipitation year?
- Black Mesa and Carrizo monthly means most closely correlated 


### Raw daily time series

```{r, include = FALSE}
# read in the data:

carrizo <- read_csv("Carrizo/carrizo_swe_03_19.csv")

chuska <- read_csv("Chuska/chuska_swe_03_19.csv")

black_mesa <- read_csv("Black Mesa/BM_swe_03_19.csv")

```

```{r, echo = FALSE}
# graph it
source('Functions/graph_swe_daily.R')
carr_daily <- graph_swe_daily(carrizo, "Carrizo")
ch_daily <- graph_swe_daily(chuska, "Chuska")
bm_daily <- graph_swe_daily(black_mesa, "Black Mesa")

# visualize all next to each other:

grid.arrange(carr_daily, ch_daily, bm_daily, heights = c(3, 3, 3))


```


### Monthly mean time series

```{r, echo = FALSE}
source('Functions/calc_swe_metrics.R')
source('Functions/graph_swe_metric.R')

# create the monthly mean dataframes

carr_monthly <- calc_swe_metrics(carrizo, "month", mean)
ch_monthly <- calc_swe_metrics(chuska, "month", mean)
bm_monthly <- calc_swe_metrics(black_mesa, "month", mean)

# graph it
carr_mn_graph <- graph_swe_metric(carr_monthly, geom_line, "Carrizo")
ch_mn_graph <- graph_swe_metric(ch_monthly, geom_line, "Chuska")
bm_mn_graph <- graph_swe_metric(bm_monthly, geom_line, "Black Mesa")

# visualize all next to each other 

grid.arrange(carr_mn_graph, ch_mn_graph, bm_mn_graph)

# is this better visually?

  graph <- carr_monthly %>% 
    ggplot(aes(x = date, y = swe_mm)) + 
    geom_line(col = "green") +
    geom_line(data = ch_monthly, aes(x = date), col = "blue") +
    geom_line(data = bm_monthly, aes(x = date), col = "red") +
    scale_y_continuous(expand = c(0,0))+
    theme_classic() + 
    labs(
      x = "Year",
      y = "SWE (mm)",
      colour = "Region"
    ) +
    theme_classic() 
 graph 

```

Black Mesa: red
Chuska: blue
Carrizo: green

### Intra Monthly Anomalies

```{r, echo = FALSE}
source('Functions/calc_month_anom.R')
source('Functions/plot_swe_anomaly.R')

# makes the dataframe of intramonth means and anomalies
carr_mn_anom <- calc_month_anom(carr_monthly)
ch_mn_anom <- calc_month_anom(ch_monthly)
bm_mn_anom <- calc_month_anom(bm_monthly)

# plot it
carr_mn_anom_graph <- plot_swe_anomaly(carr_mn_anom, "Carrizo")
ch_mn_anom_graph <- plot_swe_anomaly(ch_mn_anom, "Chuska")
bm_mn_anom_graph <- plot_swe_anomaly(bm_mn_anom, "Black Mesa")

# visualize next to each other

grid.arrange(carr_mn_anom_graph, ch_mn_anom_graph, bm_mn_anom_graph, heights = c(3, 3, 3))


```


### Correlations of Monthly Means Among Regions

```{r, include = FALSE}
# pearson's correlation
corr_chu_bm <- cor.test(bm_monthly$swe_mm, ch_monthly$swe_mm)
corr_chu_carr <- cor.test(carr_monthly$swe_mm, ch_monthly$swe_mm)
corr_carr_bm <- cor.test(bm_monthly$swe_mm, carr_monthly$swe_mm)

# linear models
corr_chu_bm_lm <- lm(bm_monthly$swe_mm ~ ch_monthly$swe_mm)
  chu_bm_r2 <- summary(corr_chu_bm_lm)$r.squared

corr_chu_carr_lm <- lm(carr_monthly$swe_mm ~ ch_monthly$swe_mm)
  chu_carr_r2 <- summary(corr_chu_carr_lm)$r.squared 

corr_carr_bm_lm <- lm(bm_monthly$swe_mm ~ carr_monthly$swe_mm)
  carr_bm_r2 <- summary(corr_carr_bm_lm)$r.squared 
  
# quick graph


```

Chuska vs Black Mesa R squared: `r chu_bm_r2`

Chuska vs Carrizo R squared: `r chu_carr_r2`

Black Mesa vs Carrizo R squared: `r carr_bm_r2`



## Appendix

Just to view these a bit better:

```{r, echo = FALSE}
carr_mn_anom_graph 
ch_mn_anom_graph
bm_mn_anom_graph 


```

