---
title: "SWE Analyses"
output: html_document
Author: AnnaClaire Marley
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the packages:

```{r, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)

```


# Carrizo

- first I need to join all the different date ranges together into one big file

```{r, message=FALSE}
# read in the files & clean up
car_03_09 <- read_csv("Carrizo/03-09-swe.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_carrizo_2003_10_01_to_2009_10_01)

car_09_15 <- read_csv("Carrizo/09-15-swe.csv")%>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_carrizo_2009_10_02_to_2015_10_01)

car_15_19 <- read_csv("Carrizo/15-19-swe.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_carrizo_2015_10_02_to_2019_06_01)

# now join them together to make one big file

carrizo_bind_most <- rbind(car_03_09, car_09_15)
carrizo_swe_03_19 <- rbind(carrizo_bind_most, car_15_19)

# export csv
write.csv(carrizo_swe_03_19, "carrizo_swe_03_19.csv", row.names = FALSE)

```

Initial data exploration:
```{r, message=FALSE}

carrizo_swe_graph <- carrizo_swe_03_19 %>% 
  ggplot(aes(x = date_time, y = SWE_mm)) +
  geom_line() +
  scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "SWE (mm)",
    title = "Carrizo Raw SWE"
  ) +
  theme_classic() 

carrizo_swe_graph

```

Annual Max SWE

```{r, message = FALSE}

# dataframe of annual maximum SWE
carr_swe_yr_max <- carrizo_swe_03_19 %>% 
  group_by(year = floor_date(date_time, "year"))  %>% 
  summarize(max_swe = max(SWE_mm))

# graph it:
carr_yr_graph <- carr_swe_yr_max %>% 
  ggplot(aes(x = year, y = max_swe)) + 
  geom_line() +
  scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "SWE (mm)",
    title = "Carrizo Yearly Max SWE"
  ) +
  theme_classic() 

carr_yr_graph


```


Weekly Max SWE

```{r, message = FALSE}
# weekly max dataframe
carr_wk_max <- carrizo_swe_03_19 %>%
  group_by(year_week = floor_date(date_time, "1 week")) %>% 
  summarize(max_swe = max(SWE_mm)) 

# weekly max graph
carr_wk_max_graph <- carr_wk_max %>% 
  ggplot(aes(x = year_week, y = max_swe)) + 
  geom_line() +
   scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "SWE (mm)",
    title = "Carrizo Weekly Max SWE"
  ) +
  theme_classic()

carr_wk_max_graph
```

Weekly mean SWE

```{r, message = FALSE}
# weekly mean dataframe
carr_wk_mean <- carrizo_swe_03_19 %>%
  group_by(year_week = floor_date(date_time, "1 week")) %>% 
  summarize(mean_swe = mean(SWE_mm)) 

#weekly mean graph
carr_wk_mean_graph <- carr_wk_mean %>% 
  ggplot(aes(x = year_week, y = mean_swe)) + 
  geom_line() +
   scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "SWE (mm)",
    title = "Carrizo Weekly Mean SWE"
  ) + 
  theme_classic()

carr_wk_mean_graph


```

Monthly mean
```{r, message = FALSE}
# monthly mean dataframe
carr_mn_mean <- carrizo_swe_03_19 %>%
  group_by(year_month = floor_date(date_time, "month")) %>% 
  summarize(mean_swe = mean(SWE_mm)) 

#weekly mean graph
carr_mn_mean_graph <- carr_mn_mean %>% 
  ggplot(aes(x = year_month, y = mean_swe)) + 
  geom_line() +
   scale_y_continuous(expand = c(0,0)) +
  labs(
    x = "Year",
    y = "SWE (mm)",
    title = "Carrizo Monthly Mean SWE"
  ) +
  theme_classic()

carr_mn_mean_graph

```

March of each year- need to work on this
```{r, message = FALSE}
# monthly max dataframe
carr_test <- carrizo_swe_03_19 %>%
  mutate(month = format(date_time,"%m")) %>% 
  filter(month == "03") %>% 
  group_by(year_week = floor_date(date_time, "1 week"))

#March 
carr_march_graph <- carr_test %>% 
  ggplot(aes(x = date_time, y = SWE_mm)) + 
  geom_point() +
  labs(
    x = "Year",
    y = "SWE (mm)"
  ) +
  theme_classic()

carr_march_graph

```



# Chuska

read in the files to make one file
```{r}

ch_03_05 <- read_csv("Chuska/ch_03-05_swe.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_chuska_2003_10_01_to_2005_10_01)


ch_05_09 <- read_csv("Chuska/ch_swe_05-09.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm =mm_swe_snodas_at_chuska_2005_10_02_to_2009_10_01)

ch_09_13 <- read_csv("Chuska/ch_swe_09-13.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_chuska_2009_10_02_to_2013_10_01)

ch_13_16 <- read_csv("Chuska/ch_swe_13-16.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_chuska_2013_10_02_to_2016_10_01)

ch_16_19 <- read_csv("Chuska/ch_swe_16-19.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_chuska_2016_10_02_to_2019_06_01)

# make one big dataset
chuska_swe_03_19 <- rbind(ch_03_05, ch_05_09, ch_09_13, ch_13_16, ch_16_19)

# write it as a csv
write.csv(chuska_swe_03_19, "chuska_swe_03_19.csv", row.names = FALSE)

```


Read data in:
```{r, message = FALSE}

# insert your csv here 
NN_region <- read_csv("Chuska/chuska_swe_03_19.csv")

```

### Graph the daily SWE data
```{r}
source('Functions/graph_swe_daily.R') # this command just sources the function

graph_swe_daily(NN_region)

```

### Get dataframe and graph for: Monthly **max** SWE
```{r, message = FALSE}
source('Functions/calc_max_mn_swe.R')

calc_max_mn_swe(NN_region)

```

### Get dataframe and graph: Monthly **mean** SWE
```{r, message = FALSE}
source('Functions/calc_av_mn_swe.R')

calc_av_mn_swe(NN_region)

```

### Get dataframe and graph: Weekly **max** SWE
```{r, message = FALSE}
source('Functions/calc_max_wk_swe.R')

calc_max_wk_swe(NN_region)

```

### Get dataframe and graph: Weekly **mean** SWE
```{r, message = FALSE}
source('~/github/CAARE_2019/Functions/calc_av_wk_swe.R')

calc_av_wk_swe(NN_region)

```


#  Black Mesa

Turn all the little dataframes into big ones
```{r}

BM_03_06 <- read_csv("Black Mesa/BM_swe_03-06.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_black_mesa_2003_10_01_to_2006_10_01)

BM_06_09 <- read_csv("Black Mesa/BM_swe_06-09.csv") %>% 
  clean_names() %>% 
  rename(SWE_mm = mm_swe_snodas_at_black_mesa_2006_10_02_to_2009_10_01)

BM_09_12 <- read_csv("Black Mesa/BM_swe_09-12.csv") %>% 
  clean_names() %>%
  rename(SWE_mm = mm_swe_snodas_at_black_mesa_2009_10_02_to_2012_10_01)

BM_12_15 <- read_csv("Black Mesa/BM_swe_12-15.csv") %>% 
  clean_names() %>%
  rename(SWE_mm = mm_swe_snodas_at_black_mesa_2012_10_02_to_2015_10_01)

BM_15_18 <- read_csv("Black Mesa/BM_swe_15-18.csv") %>% 
  clean_names() %>%
  rename(SWE_mm = mm_swe_snodas_at_black_mesa_2015_10_02_to_2018_10_01)

BM_18_19 <- read_csv("Black Mesa/BM_swe_18-19.csv") %>% 
  clean_names() %>%
  rename(SWE_mm = mm_swe_snodas_at_black_mesa_2018_10_02_to_2019_06_01)

# make one dataframe

BM_swe_03_19 <- rbind(BM_03_06, BM_06_09, BM_09_12, BM_12_15, BM_15_18, BM_18_19)

write.csv(BM_swe_03_19, "BM_swe_03_19.csv", row.names = FALSE)

```

