---
title: "SWE Functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Instructions** 

- This Rmarkdown document will allow you to run basic analyses and make graphs of SWE max and mean weekly, monthly, and annually. 
- If you want to save the function in your environment make sure to call it something.
- Make sure your raw data date column is named "date_time" and SWE is "SWE_mm"

Load packages
```{r, include = FALSE}

library(tidyverse)
library(janitor)
library(lubridate)

```

Read your data in:
```{r, message = FALSE}

# insert your csv here 
NN_region <- read_csv("carrizo_swe_03_19.csv")

```

### Graph the daily SWE data
```{r}
source('Functions/graph_swe_daily.R') # this command just sources the function

graph_swe_daily(NN_region)

```

### Get dataframe and graph for: Annual **Max** SWE
```{r, message = FALSE}
source('Functions/calc_max_yr_swe.R')

calc_max_yr_swe(NN_region)

# if you want just the dataframe or just the plot, put $ sign at end to just call and name that
# example <- calc_max_yr_swe(carrizo_swe_03_19)$swe_yr_max

```

### Get dataframe and graph for: Monthly **max** SWE
```{r, message = FALSE}
source('Functions/calc_max_mn_swe.R')

calc_max_mn_swe(NN_region)

```

### Get dataframe and graph: Monthly **mean** SWE
```{r, message = FALSE}
source('Functions/calc_av_mn_swe.R')

calc_av_mn_swe(NN_region)

```

### Get dataframe and graph: Weekly **max** SWE
```{r, message = FALSE}
source('Functions/calc_max_wk_swe.R')

calc_max_wk_swe(NN_region)

```

### Get dataframe and graph: Weekly **mean** SWE
```{r, message = FALSE}
source('~/github/CAARE_2019/Functions/calc_av_wk_swe.R')

calc_av_wk_swe(NN_region)

```

