---
title: "SWE & Snow Depth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
source("Functions/util.R")
source('Functions/graphing.R')
```


### This rmarkdown shows some of the relationships between SWE and snow depth at each location. Analyses conducted include:

- Correlation between SWE and Snow Depth with the colors grouped by months
- linear relationship between Swe and snow depth for each month
- graph of the slopes of the monthly linear relationship


## General Patterns

#### Correlation between SWE and Snow Depth (November - April Water Year)

- SWE and snow depth are highly correlated across regions (>0.9)
- Generally, as snow depth and swe values get larger, they become less closely correlated

#### Monthly linear relationships between SWE and Snow Depth 

- SWE is a function of depth and density, so SWE/depth = bulk density
- So through the water year, we see generally increasing density at all locations November to March. 
  - However, in April, there is sometimes a drop in density depending on location - this may simply be because most of the snow has melted

```{r, include = FALSE}
# read in the daily data files
car_swe <- read_csv("winter_wy/car_swe_wy_03_19.csv")
car_dpth <- read_csv("winter_wy/car_dpth_wy_03_19.csv")

# combine to make a dataframe of swe and dpth daily data
car_swe_dpth <- merge_snow(car_swe, car_dpth)

# monthly means correlation
car_swe_av <- calc_swe_metrics(car_swe, "month", mean)
car_dpth_av <- calc_depth_metrics(car_dpth, "month", mean)

```

```{r, include = FALSE}
calc_swe_metrics(car_swe_dpth, "year", max)
```


### Carrizo 

- Generally, January, February, and March swe and depth values are less correlated than other months


```{r, include = FALSE}
# average swe - depth metric (from swe/depth)
car_av_ratio <- car_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

car_av_snow_ratio <- mean(car_av_ratio$mean_ratio)
  

```


```{r, echo=FALSE}
# daily data linear regression:

car_swe_dpth_lm <- lm(car_swe_dpth$swe_mm ~ car_swe_dpth$depth_mm)
car_swedpth_r2 <- summary(car_swe_dpth_lm)$adj.r.squared

```

```{r, echo = FALSE}
# graph it 

car_daily_cor_swe_dpth <- plot_cor(car_swe_dpth, car_swedpth_r2, region = "Carrizo")
 car_daily_cor_swe_dpth

```

Each month has it's own line:

```{r,echo= FALSE}

plot_cor_month(car_swe_dpth, car_swedpth_r2, region = "Carrizo")



```


Slope of the line of the linear relationship of each month's swe and snow depth
- almost linearly increasing slop through the water year until April, where the slope decreases back down to almost the February slope


```{r, echo = FALSE}
# equations of the lines:

car_mnth_equ <- cor_equation(car_swe_dpth)
plot_slope_swedpth(car_mnth_equ, title = "Carrizo")

```

# {.tabset .tabset-fade}

## Chuska 

- Generally, January, February, March, amd some April swe and depth values are less correlated than other months

```{r, include=FALSE}
# read in the files
ch_swe <- read_csv("winter_wy/ch_swe_wy_03_19.csv")
ch_dpth <- read_csv("winter_wy/ch_dpth_wy_03_19.csv")

# merge them together 
ch_swe_dpth <- merge_snow(ch_swe, ch_dpth)

```


```{r, include = FALSE}
# average swe - depth metric (from swe/depth)
ch_av_ratio <- ch_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

ch_av_snow_ratio <- mean(ch_av_ratio$mean_ratio)
  


```


```{r, echo = FALSE}
# linear regression
ch_swe_v_depth_lm <- lm(ch_swe_dpth$swe_mm ~ ch_swe_dpth$depth_mm)
ch_swedepth_r2 <- summary(ch_swe_v_depth_lm)$adj.r.squared 


# plot it
ch_swe_v_depth_graph <- plot_cor(ch_swe_dpth, ch_swedepth_r2, region = "Chuska")
ch_swe_v_depth_graph 
```



Each month has it's own line:

```{r, echo = FALSE}

plot_cor_month(ch_swe_dpth, car_swedpth_r2, region = "") + 
  geom_point(alpha = 0.2) + 
  scale_color_discrete(labels= c("Jan", "Feb", "Mar", "Apr", "Nov", "Dec")) 


```

- almost perfect linear increase in the slope of the relationship between swe and depth through time
- adjusted r squared values for all lines > 0.9
```{r, echo = FALSE}
# equations of the lines:

ch_mnth_equ <- cor_equation(ch_swe_dpth)  

plot_slope_swedpth(ch_mnth_equ, title = "Chuska")


```

# {.tabset .tabset-fade}

## Black Mesa 

- Generally, January, February, and March swe and depth values are less correlated than other months

```{r, include=FALSE}
# read in the files
bm_swe <- read_csv("winter_wy/bm_swe_wy_03_19.csv")
bm_dpth <- read_csv("winter_wy/bm_dpth_wy_03_19.csv")

# merge them together 
bm_swe_dpth <- merge_snow(bm_swe, bm_dpth)

```


```{r, include = FALSE}
# average density (from swe/depth)
bm_av_ratio <- bm_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

bm_av_snow_ratio <- mean(bm_av_ratio$mean_ratio)
  


```


```{r, echo = FALSE}
# linear regression
bm_swe_v_depth_lm <- lm(bm_swe_dpth$swe_mm ~ bm_swe_dpth$depth_mm)
bm_swedepth_r2 <- summary(bm_swe_v_depth_lm)$adj.r.squared 


# plot it
bm_swe_v_depth_graph <- plot_cor(bm_swe_dpth, bm_swedepth_r2, region = "Black Mesa")
bm_swe_v_depth_graph
```


Each month has it's own line:

```{r, echo = FALSE}

plot_cor_month(bm_swe_dpth, car_swedpth_r2, region = "Black Mesa")

```

- almost linearly increasing slop through the water year until April, where the slope decreases back down to almost the November slope

```{r, echo = FALSE}
# equations of the lines:

bm_mnth_equ <- cor_equation(bm_swe_dpth)
plot_slope_swedpth(bm_mnth_equ, title = "Black Mesa")
```

# {.tabset .tabset-fade}

## Navajo Mt 

- March, January, and some december most often less correlated
```{r, include=FALSE}
# read in the files
nm_swe <- read_csv("winter_wy/nm_swe_wy_03_19.csv")
nm_dpth <- read_csv("winter_wy/nm_dpth_wy_03_19.csv")

# merge them together 
nm_swe_dpth <- merge_snow(nm_swe, nm_dpth)

```


```{r, include = FALSE}
# average density (from swe/depth)
nm_av_ratio <- nm_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

nm_av_snow_ratio <- mean(nm_av_ratio$mean_ratio)
  


```


```{r, echo = FALSE}
# linear regression
nm_swe_v_depth_lm <- lm(nm_swe_dpth$swe_mm ~ nm_swe_dpth$depth_mm)
nm_swedepth_r2 <- summary(nm_swe_v_depth_lm)$adj.r.squared 


# plot it
nm_swe_v_depth_graph <- plot_cor(nm_swe_dpth, nm_swedepth_r2, region = "Navajo Mt")
nm_swe_v_depth_graph
```


Each month has it's own line:

```{r, echo = FALSE}

plot_cor_month(nm_swe_dpth, car_swedpth_r2, region = "Navajo Mountain")


```

- Seems to be the only location where there is not really a clear linear increase in the slope of the relationship between swe and snow depth

```{r, echo = FALSE}

# equations of the lines:

nm_mnth_equ <- cor_equation(nm_swe_dpth)
plot_slope_swedpth(nm_mnth_equ, title = "Navajo Mt")

```

# {.tabset .tabset-fade}

## Mt Powell 

- Jn, Feb and March seem to be outliers

```{r, include=FALSE}
# read in the files
mp_swe <- read_csv("winter_wy/mp_swe_wy_03_19.csv")
mp_dpth <- read_csv("winter_wy/mp_dpth_wy_03_19.csv")

# merge them together 
mp_swe_dpth <- merge_snow(mp_swe, mp_dpth)

```


```{r, include = FALSE}
# average density (from swe/depth)
mp_av_ratio <- mp_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

mp_av_snow_ratio <- mean(mp_av_ratio$mean_ratio)
  


```


```{r, echo = FALSE}
# linear regression
mp_swe_v_depth_lm <- lm(mp_swe_dpth$swe_mm ~ mp_swe_dpth$depth_mm)
mp_swedepth_r2 <- summary(mp_swe_v_depth_lm)$adj.r.squared 


# plot it
mp_swe_v_depth_graph <- plot_cor(mp_swe_dpth, mp_swedepth_r2, region = "Mt Powell")
mp_swe_v_depth_graph
```


Each month has it's own line:

```{r, echo = FALSE}

plot_cor_month(mp_swe_dpth, car_swedpth_r2, region = "Mt Powell")



```

- almost linearly increasing slop through the water year until April, where the slope decreases back down to almost the November slope

```{r, echo = FALSE}

# equations of the lines:

mp_mnth_equ <- cor_equation(mp_swe_dpth)
plot_slope_swedpth(mp_mnth_equ, title = "Mt Powell")

```

# {.tabset .tabset-fade}

## Defiance Plateau 

- March seems to be least correlated

```{r, include=FALSE}
# read in the files
dp_swe <- read_csv("winter_wy/dp_swe_wy_03_19.csv")
dp_dpth <- read_csv("winter_wy/dp_dpth_wy_03_19.csv")

# merge them together 
dp_swe_dpth <- merge_snow(dp_swe, dp_dpth)

```


```{r, include = FALSE}
# average density (from swe/depth)
dp_av_ratio <- dp_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

dp_av_snow_ratio <- mean(dp_av_ratio$mean_ratio)
  


```


```{r, echo = FALSE}
# linear regression
dp_swe_v_depth_lm <- lm(dp_swe_dpth$swe_mm ~ dp_swe_dpth$depth_mm)
dp_swedepth_r2 <- summary(dp_swe_v_depth_lm)$adj.r.squared 


# plot it
dp_swe_v_depth_graph <- plot_cor(dp_swe_dpth, dp_swedepth_r2, region = "Defiance Plateau")
dp_swe_v_depth_graph
```


Each month has it's own line:

```{r, echo = FALSE}

plot_cor_month(dp_swe_dpth, car_swedpth_r2, region = "Defiance Plateau")




```

- almost linearly increasing slop through the water year until April, where the slope decreases back down to almost the January slope

```{r, echo = FALSE}

# equations of the lines:

dp_mnth_equ <- cor_equation(dp_swe_dpth)
plot_slope_swedpth(dp_mnth_equ, title = "Defiance Plateau")

```

# {.tabset .tabset-fade}

## SWE-depth ratio 

- annual averaged metric 

- **Carrizo**: `r round(car_av_snow_ratio, 3)`

- **Chuska**: `r round(ch_av_snow_ratio, 3)`

- **Black Mesa**: `r round(bm_av_snow_ratio, 3)`
.
### SWE-depth ratio (= density) monthly average: 

- Generally, density increases through the water year (starting at november). This corroborates the trend seen in the correlation graph where depth and swe becomes less correlated as the water year progresses

```{r, echo = FALSE}

# create a dataframe of the density of each lcoation
all_density <- merge(bm_av_ratio, ch_av_ratio, by = "month") %>% 
  merge(car_av_ratio, by = "month") %>% 
  rename(black_mesa = mean_ratio.x, chuska = mean_ratio.y, carrizo = mean_ratio) %>% 
  mutate(black_mesa = round(black_mesa, 3))  %>% 
  mutate(carrizo = round(carrizo, 3))  %>% 
  mutate(chuska = round(chuska, 3)) %>% 
  merge(mp_av_ratio, by = "month") %>% 
  rename(mt_powell = mean_ratio) %>% 
  merge(nm_av_ratio, by = "month") %>% 
  rename(navajo_mt = mean_ratio) %>% 
  merge(dp_av_ratio, by = "month") %>% 
  rename(defiance_plateau = mean_ratio)

all_density_tidy <- all_density %>% 
  mutate(month_n = c("January", "February", "March", "April", "November", "December")) %>% 
  mutate(month_n = factor(month_n, levels = c("November", "December", "January", "February", "March", "April"))) %>% 
  mutate(month = factor(month, levels = c(11, 12, 1, 2, 3, 4))) %>% 
  melt(id.vars= c("month", "month_n")) %>% 
  rename(region = variable) %>% 
  rename(density = value)


# graph it

all_density_graph <- all_density_tidy %>% 
  ggplot(aes(x = month_n, y = density, group = region, color = region)) +
  geom_line() +
  theme_classic() +
  labs(x = "Month",
       y = "Density",
       color = "Region") +
  scale_color_discrete(labels = c("Black Mesa", "Chuska", "Carrizo", "Mt Powell", "Navajo Mt", "Defiance Plateau")) +
  scale_x_discrete(expand = c(0,0))

all_density_graph

```

```{r, fig.width=20,fig.height=5, echo = FALSE}

car_wk_mean_swe <- calc_swe_metrics(car_swe_dpth, "week", mean) 
ch_wk_mean_swe <- calc_swe_metrics(ch_swe_dpth, "week", mean) 
nm_wk_mean_swe <- calc_swe_metrics(nm_swe_dpth, "week", mean) 
mp_wk_mean_swe <- calc_swe_metrics(mp_swe_dpth, "week", mean) 
dp_wk_mean_swe <- calc_swe_metrics(dp_swe_dpth, "week", mean) 
bm_wk_mean_swe <- calc_swe_metrics(bm_swe_dpth, "week", mean) 

graph_with_wateryear(car_wk_mean_swe, type = geom_col,  variable = "swe_mm", ylab = "SWE (mm)")



```



### SWE and Snow depth (November - April Water Year) 

- Generally, snow depth is much greater than SWE, but follows the same general patter

```{r, fig.width=20,fig.height=5, echo = FALSE, warning=FALSE, message=FALSE}
# carrizo 
graph_with_wateryear(car_swe_dpth, title = "Carrizo") +
  geom_line(aes(y = depth_mm), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)")) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue"))
  

```


```{r, fig.width=20,fig.height=5, echo = FALSE, warning=FALSE, message=FALSE}
# chuska
graph_with_wateryear(ch_swe_dpth, title = "Chuska") +
  geom_line(aes(y = depth_mm), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)")) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue"))
  

```

```{r, fig.width=20,fig.height=5,echo = FALSE, warning=FALSE, message=FALSE}
# black mesa
graph_with_wateryear(bm_swe_dpth, title = "Black Mesa") +
  geom_line(aes(y = depth_mm), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)")) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue")) 
  

```

### Chuska SWE vs depth monthly anomaly

```{r, echo = FALSE}
## Monthly anom
# create chuska swe anomaly 

ch_swe_anom <- calc_month_anom(calc_swe_metrics(ch_swe, "month", mean)) %>% 
  rename(swe_anomaly = anomaly) %>% 
  rename(swe_anom_perc = anomaly_perc)

# chuska depth anomaly

ch_depth_anom <- calc_mnth_depth_anom(calc_depth_metrics(ch_dpth, "month", mean)) %>% 
  rename(depth_anom_perc = anomaly_perc)

# merge together 

ch_swe_dpth_anom <- merge(ch_depth_anom, ch_swe_anom, by = "date") %>% 
  select(date, waterYear.x, anomaly, swe_anom_perc, depth_anom_perc, sign.x, swe_anomaly) %>% 
  rename(waterYear = waterYear.x)



```

```{r, echo = FALSE}

# create graph 
      # first make variables for ease of graphing  
      startWaterYear = min(ch_swe_dpth_anom$waterYear)
      endWaterYear = max(ch_swe_dpth_anom$waterYear)
      # Create sequence of real dates, on which to place the water year lables
      breakDates = seq.Date(from = ymd(paste(startWaterYear,1,1, sep='-')), length.out = endWaterYear - startWaterYear + 1, by = "1 year")
      
  ggplot(ch_swe_dpth_anom, aes(x = date)) +
  geom_col(aes(y = anomaly, fill = sign.x), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_date(breaks= breakDates, labels=seq(startWaterYear,endWaterYear,by=1)) +
  labs(
    x = "Water Year",
    y = "Snow Depth Anomaly (mm)"
  ) +
  theme_classic() +
  geom_line(aes(y = swe_anomaly), color = "#1219cc") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "SWE Anomaly (mm)"), expand = c(0,0)) +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
        axis.text.y.right = element_text(color = "#1219cc"),
        axis.title.y.right = element_text(color = "#1219cc")) 

```

```{r, echo = FALSE}
### Percent difference from average###
# create graph 
      # first make variables for ease of graphing  
      startWaterYear = min(ch_swe_dpth_anom$waterYear)
      endWaterYear = max(ch_swe_dpth_anom$waterYear)
      # Create sequence of real dates, on which to place the water year lables
      breakDates = seq.Date(from = ymd(paste(startWaterYear,1,1, sep='-')), length.out = endWaterYear - startWaterYear + 1, by = "1 year")
      
  ggplot(ch_swe_dpth_anom, aes(x = date)) +
  geom_col(aes(y = depth_anom_perc, fill = sign.x), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_date(breaks= breakDates, labels=seq(startWaterYear,endWaterYear,by=1)) +
  labs(
    x = "Water Year",
    y = "Snow Depth % Difference from Average"
  ) +
  theme_classic() +
  geom_line(aes(y = swe_anom_perc), color = "#1219cc") +
  scale_y_continuous(sec.axis = sec_axis(~., name = "SWE % Difference from Average"), 
                     expand = c(0,0)) +
  theme(axis.line.y.right = element_line(color = "#1219cc"),
        axis.text.y.right = element_text(color = "#1219cc"),
        axis.title.y.right = element_text(color = "#1219cc")) 

```

```{r, echo = FALSE}

ch_nov_anom <- ch_swe_anom %>% 
  filter(month == "11") %>% 
  ggplot(aes(x = waterYear, y = swe_anom_perc)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_continuous(breaks = seq(2004,2019, by = 3)) +
  labs(
    x = "Water Year",
    y = "SWE % Difference from Average",
    title = "November"
  ) +
  theme_classic()



ch_dec_anom <- ch_swe_anom %>% 
  filter(month == "12") %>% 
  ggplot(aes(x = waterYear, y = swe_anom_perc)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_continuous(breaks = seq(2004,2019, by = 3)) +
  labs(
    x = "Water Year",
    y = "SWE % Difference from Average",
    title = "December"
  ) +
  theme_classic()



# just january anomalies

ch_jan_anom <- ch_swe_anom %>% 
  filter(month == "1") %>% 
  ggplot(aes(x = waterYear, y = swe_anom_perc)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_continuous(breaks = seq(2004,2019, by = 3)) +
  labs(
    x = "Water Year",
    y = "SWE % Difference from Average",
    title = "January"
  ) +
  theme_classic()



# just feb anomalies
ch_feb_anom <- ch_swe_anom %>% 
  filter(month == "2") %>% 
  ggplot(aes(x = waterYear, y = swe_anom_perc)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_continuous(breaks = seq(2004,2019, by = 3)) +
  labs(
    x = "Water Year",
    y = "SWE % Difference from Average",
    title = "February"
  ) +
  theme_classic()


# just march anomalies
ch_mar_anom <- ch_swe_anom %>% 
  filter(month == "3") %>% 
  ggplot(aes(x = waterYear, y = swe_anom_perc)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_continuous(breaks = seq(2004,2019, by = 3)) +
  labs(
    x = "Water Year",
    y = "SWE % Difference from Average",
    title = "March"
  ) +
  theme_classic()



# just march anomalies
ch_apr_anom <- ch_swe_anom %>% 
  filter(month == "4") %>% 
  ggplot(aes(x = waterYear, y = swe_anom_perc)) +
  geom_col(aes(fill = sign), show.legend = FALSE) +
  scale_fill_manual(values = c("negative" = "#df5e3d", "positive" = "#a6c39d")) +
  scale_x_continuous(breaks = seq(2004,2019, by = 3)) +
  scale_y_continuous(breaks = seq(-100,800, by = 100)) +
  labs(
    x = "Water Year",
    y = "SWE % Difference from Average",
    title = "April"
  ) +
  theme_classic()


```

```{r, fig.width= 10, echo = FALSE}

grid.arrange(ch_nov_anom, ch_dec_anom, ch_jan_anom, ch_feb_anom, ch_mar_anom, ch_apr_anom, ncol = 3)

```


### Depth scaled (depth*0.2) by swe/depth metric

- shows how well correlated SWE and snow depth are

```{r, fig.width=20,fig.height=5, echo = FALSE, message = FALSE}

graph_with_wateryear(bm_swe_dpth, title = "Black Mesa", type = geom_col) +
  geom_line(aes(y = depth_mm*0.2), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)"), expand = c(0,0)) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue")) 
  

```



