---
title: "SWE & Snow Depth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid)
library(dataRetrieval)
library(reshape2)
source("Functions/util.R")
source('Functions/graphing.R')
```



### Correlation between SWE and Snow Depth (November - April Water Year)

- SWE and snow depth are highly correlated across regions (>0.9)
- Generally, as snow depth and swe values get larger, they become less closely correlated



```{r, include = FALSE}
# read in the daily data files
car_swe <- read_csv("winter_wy/car_swe_wy_03_19.csv")
car_dpth <- read_csv("winter_wy/car_dpth_wy_03_19.csv")

# combine to make a dataframe of swe and dpth daily data
car_swe_dpth <- merge_snow(car_swe, car_dpth)

# monthly means correlation
car_swe_av <- calc_swe_metrics(car_swe, "month", mean)
car_dpth_av <- calc_depth_metrics(car_dpth, "month", mean)

```

### Carrizo

- Generally, January, February, and March swe and depth values are less correlated than other months


```{r, include = FALSE}
# average swe - depth metric (from swe/depth)
car_av_ratio <- car_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

car_av_snow_ratio <- mean(car_av_ratio$mean_ratio)
  

```


```{r, echo=FALSE}
# daily data linear regression:

car_swe_dpth_lm <- lm(car_swe_dpth$swe_mm ~ car_swe_dpth$depth_mm)
car_swedpth_r2 <- summary(car_swe_dpth_lm)$adj.r.squared

```

```{r, echo = FALSE}
# graph it 

car_daily_cor_swe_dpth <- plot_cor(car_swe_dpth, car_swedpth_r2, region = "Carrizo")
 car_daily_cor_swe_dpth

```


### Chuska 

- Generally, January, February, March, amd some April swe and depth values are less correlated than other months

```{r, include=FALSE}
# read in the files
ch_swe <- read_csv("winter_wy/ch_swe_wy_03_19.csv")
ch_dpth <- read_csv("winter_wy/ch_dpth_wy_03_19.csv")

# merge them together 
ch_swe_dpth <- merge_snow(ch_swe, ch_dpth)

```


```{r, include = FALSE}
# average swe - depth metric (from swe/depth)
ch_av_ratio <- ch_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

ch_av_snow_ratio <- mean(ch_av_ratio$mean_ratio)
  


```




```{r, echo = FALSE}
# linear regression
ch_swe_v_depth_lm <- lm(ch_swe_dpth$swe_mm ~ ch_swe_dpth$depth_mm)
ch_swedepth_r2 <- summary(ch_swe_v_depth_lm)$adj.r.squared 


# plot it
ch_swe_v_depth_graph <- plot_cor(ch_swe_dpth, ch_swedepth_r2, region = "Chuska")
ch_swe_v_depth_graph 
```

### Black Mesa

- Generally, January, February, and March swe and depth values are less correlated than other months

```{r, include=FALSE}
# read in the files
bm_swe <- read_csv("winter_wy/bm_swe_wy_03_19.csv")
bm_dpth <- read_csv("winter_wy/bm_dpth_wy_03_19.csv")

# merge them together 
bm_swe_dpth <- merge_snow(bm_swe, bm_dpth)

```


```{r, include = FALSE}
# average swe - depth metric (from swe/depth)
bm_av_ratio <- bm_swe_dpth %>% 
  filter(snow_ratio != "NaN") %>% 
  filter(snow_ratio != "Inf") %>% 
  group_by(month) %>% 
  summarize(
    mean_ratio = mean(snow_ratio)
  )
  

bm_av_snow_ratio <- mean(bm_av_ratio$mean_ratio)
  


```


```{r, echo = FALSE}
# linear regression
bm_swe_v_depth_lm <- lm(bm_swe_dpth$swe_mm ~ bm_swe_dpth$depth_mm)
bm_swedepth_r2 <- summary(bm_swe_v_depth_lm)$adj.r.squared 


# plot it
bm_swe_v_depth_graph <- plot_cor(bm_swe_dpth, bm_swedepth_r2, region = "Black Mesa")
bm_swe_v_depth_graph
```

### SWE-depth ratio - annual averaged metric 

- **Carrizo**: `r round(car_av_snow_ratio, 3)`

- **Chuska**: `r round(ch_av_snow_ratio, 3)`

- **Black Mesa**: `r round(bm_av_snow_ratio, 3)`

### SWE-depth ratio monthly average:

- Generally, the ratio increases through the water year (starting at november). This corroborates the trend seen in the correlation graph where depth and swe becomes less correlated as the water year progresses

```{r, echo = FALSE}

all_swe_dpth <- merge(bm_av_ratio, ch_av_ratio, by = "month") %>% 
  merge(car_av_ratio, by = "month") %>% 
  rename(black_mesa = mean_ratio.x, chuska = mean_ratio.y, carrizo = mean_ratio) %>% 
  mutate(black_mesa = round(black_mesa, 3))  %>% 
  mutate(carrizo = round(carrizo, 3))  %>% 
  mutate(chuska = round(chuska, 3)) 
all_swe_dpth

  

```



### SWE and Snow depth (November - April Water Year)

- Generally, snow depth is much greater than SWE, but follows the same general patter

```{r, fig.width=20,fig.height=5, echo = FALSE, warning=FALSE, message=FALSE}
# carrizo 
graph_with_wateryear(car_swe_dpth, title = "Carrizo") +
  geom_line(aes(y = depth_mm), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)")) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue"))
  

```


```{r, fig.width=20,fig.height=5, echo = FALSE, warning=FALSE, message=FALSE}
# carrizo 
graph_with_wateryear(ch_swe_dpth, title = "Chuska") +
  geom_line(aes(y = depth_mm), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)")) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue"))
  

```

```{r, fig.width=20,fig.height=5,echo = FALSE, warning=FALSE, message=FALSE}
# carrizo 
graph_with_wateryear(bm_swe_dpth, title = "Black Mesa") +
  geom_line(aes(y = depth_mm), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)")) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue")) 
  

```

### Depth scaled (depth*0.2) by swe/depth metric

- shows how well correlated SWE and snow depth are

```{r, fig.width=20,fig.height=5, echo = FALSE, message = FALSE}

graph_with_wateryear(bm_swe_dpth, title = "Black Mesa", type = geom_col) +
  geom_line(aes(y = depth_mm*0.2), color = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Snow Depth (mm)"), expand = c(0,0)) +
  theme(axis.line.y.right = element_line(color = "blue"),
        axis.text.y.right = element_text(color = "blue"),
        axis.title.y.right = element_text(color = "blue")) 
  

```



