---
title: "SWE Functions"
output: html_document
Author: AnnaClaire Marley
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Instructions** 

- This Rmarkdown document will allow you to run basic analyses and make graphs of SWE max and mean weekly, monthly, and annually. 
- If you want to save the function in your environment make sure to call it something.
- Make sure your raw data date column is named "date_time" and SWE is "SWE_mm"

Load packages
```{r, include = FALSE}

library(tidyverse)
library(janitor)
library(lubridate)

```

Read your data in:
```{r, message = FALSE}

# insert your csv here 
NN_region <- read_csv("Black Mesa/BM_swe_03_19.csv")

```

This is how you can do summary statistics on your dataset. Enter in your cleaned csv file, the time period you want, ("1 week", "month", "year") *making sure it's in quotes*, and then the summary statistic (ie mean, max, min)
```{r}
source('Functions/calc_swe_metrics.R')

test <- calc_swe_metrics(NN_region, "month", mean)

```

Now graph and visualize what you found. Enter in the dataframe you created above, and then what kind of plot you want to make (ie geom_line, geom_point, geom_col etc)
```{r}
source('Functions/graph_swe_metric.R')

graph_swe_metric(test, geom_line)
```

Plot anomalies made from average over the year:

```{r}


BM_monthly <- calc_swe_metrics(NN_region, "month", mean)

source('Functions/plot_swe_anomaly.R')

plot_swe_anomaly(BM_monthly)

```


Plot anomalies from average for each month:
```{r}
source('Functions/calc_month_anom.R')

# makes the dataframe of intramonth means and anomalies
BM_intra_month <- calc_month_anom(BM_monthly)

# plots it
plot_swe_anomaly(BM_intra_month)

```

