---
title: "SWE Final Analyses"
output: html_document
Author: AnnaClaire Marley
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown is a conglomerate of all the graphs created in SWE analyses from SNODAS data collected from Climate Engine. 

Analyses shows results for SWE data from November - April of each year

```{r, include = FALSE}
# load the packages
library(tidyverse)
library(janitor)
library(lubridate)
library(gridExtra)
library(grid) 
library(dataRetrieval)


```

### Prelimary Analyses from data below:

- Chuska recieves the most snow out of the three
- 2010 is a big anomaly at all three locations, was this a heavy precipitation year?
- Black Mesa and Carrizo are most closely correlated
- Monthly medians seem pretty similar to month means


```{r, include = FALSE}

# read in the data:

carrizo <- read_csv("Carrizo/carrizo_swe_03_19.csv") 
  
chuska <- read_csv("Chuska/chuska_swe_03_19.csv")

black_mesa <- read_csv("Black Mesa/BM_swe_03_19.csv")

# clean and make it just winter data

car_wint_cal <- carrizo %>% 
  rename(date = date_time) %>% 
   filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

ch_wint_cal <- chuska %>% 
  rename(date = date_time) %>% 
   filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

bm_wint_cal <- black_mesa %>% 
  rename(date = date_time) %>% 
   filter(month(date) %in% c(11, 12, 1, 2, 3, 4))

#Changes the date column to be in water years, you don't have to do this step if you already have the dates in water year format
source('Functions/add_water_year.R')
 bm_wint_wy <- add_water_year(bm_wint_cal)
 
 car_wint_wy <- add_water_year(car_wint_cal)
 ch_wint_wy <- add_water_year(ch_wint_cal)
 
 # turn these into csv
 
 write.csv(bm_wint_wy, "Black Mesa/bm_wint_wy_04-19", row.names = FALSE)
 write.csv(car_wint_wy, "Carrizo/car_wint_wy_04-19", row.names = FALSE)
 write.csv(ch_wint_wy, "Chuska/ch_wint_wy_04-19", row.names = FALSE)

```



### Raw daily time series (November - April Calendar Year)

```{r, echo = FALSE, warning=FALSE}
# graph it
source('Functions/graph_swe_daily.R')
carr_daily <- graph_swe_daily(car_wint_cal, "Carrizo")
ch_daily <- graph_swe_daily(ch_wint_cal, "Chuska")
bm_daily <- graph_swe_daily(bm_wint_cal, "Black Mesa")

# visualize all next to each other:

grid.arrange(carr_daily, ch_daily, bm_daily, heights = c(3, 3, 3))


```

```{r, echo = FALSE, include=FALSE, error = TRUE}

## This will have to be for calendar year

source('Functions/calc_swe_metrics.R')


# create the yearly mean dataframes

carr_yr <- calc_swe_metrics(car_wint_cal, "year", mean) %>% 
  mutate(date = year(date))

ch_yr <- calc_swe_metrics(ch_wint_cal, "year", mean)%>% 
  mutate(date = year(date))

bm_yr <- calc_swe_metrics(bm_wint_cal, "year", mean)%>% 
  mutate(date = year(date))



```


## Anual average - (November-April Water Year)

```{r}
# This is for water years

source('Functions/calc_anwy_metrics.R')

carr_anwy <- calc_anwy_metrics(car_wint_wy, mean)
ch_anwy <- calc_anwy_metrics(ch_wint_wy, mean)
bm_anwy <- calc_anwy_metrics(bm_wint_wy, mean)



```

```{r}
source('Functions/graph_swe_metric.R')
# graph metrics
  #put in what type of graph you want and then the name of the region in quotes

carr_wy_graph <- graph_swe_metric(carr_anwy, "waterYear", geom_col,  "Carrizo")

ch_wy_graph <- graph_swe_metric(ch_anwy, "waterYear", geom_col, "Chuska")

bm_wy_graph <- graph_swe_metric(bm_anwy, "waterYear", geom_col, "Black Mesa")

# visualize all next to each other 

grid.arrange(carr_wy_graph, ch_wy_graph, bm_wy_graph)

```


### Monthly mean time series

```{r, echo = FALSE, error=TRUE}
source('Functions/calc_swe_metrics.R')
source('Functions/graph_swe_metric.R')

# create the monthly mean dataframes

carr_monthly <- calc_swe_metrics(car_wint_cal, "month", mean)
ch_monthly <- calc_swe_metrics(ch_wint_cal, "month", mean)
bm_monthly <- calc_swe_metrics(bm_wint_cal, "month", mean)

# graph it
carr_mn_graph <- graph_swe_metric(carr_monthly, geom_line, "Carrizo")
ch_mn_graph <- graph_swe_metric(ch_monthly, geom_line, "Chuska")
bm_mn_graph <- graph_swe_metric(bm_monthly, geom_line, "Black Mesa")

# visualize all next to each other 

grid.arrange(carr_mn_graph, ch_mn_graph, bm_mn_graph)

# is this better visually?

  graph <- carr_monthly %>% 
    ggplot(aes(x = date, y = swe_mm)) + 
    geom_line(col = "green") +
    geom_line(data = ch_monthly, aes(x = date), col = "blue") +
    geom_line(data = bm_monthly, aes(x = date), col = "red") +
    scale_y_continuous(expand = c(0,0))+
    scale_x_date(date_breaks = "2 years") +
    theme_classic() + 
    labs(
      x = "Year",
      y = "SWE (mm)",
      colour = "Region"
    ) +
    theme_classic() 
 graph 

```

Black Mesa: red
Chuska: blue
Carrizo: green

### Monthly medians

```{r, echo = FALSE, error=TRUE}
# create the monthly mean dataframes

carr_mnth_median <- calc_swe_metrics(car_wint_cal, "month", median)
ch_mnth_median <- calc_swe_metrics(ch_wint_cal, "month", median)
bm_mnth_median <- calc_swe_metrics(bm_wint_cal, "month", median)

# graph it
carr_mn_graph <- graph_swe_metric(carr_mnth_median, geom_line, "Carrizo")
ch_mn_graph <- graph_swe_metric(ch_mnth_median, geom_line, "Chuska")
bm_mn_graph <- graph_swe_metric(bm_mnth_median, geom_line, "Black Mesa")

# visualize all next to each other 

grid.arrange(carr_mn_graph, ch_mn_graph, bm_mn_graph)


```



### Intra Monthly Anomalies

```{r, echo = FALSE}
source('Functions/calc_month_anom.R')
source('Functions/plot_swe_anomaly.R')

# makes the dataframe of intramonth means and anomalies
carr_mn_anom <- calc_month_anom(carr_monthly)
ch_mn_anom <- calc_month_anom(ch_monthly)
bm_mn_anom <- calc_month_anom(bm_monthly)

# plot it
carr_mn_anom_graph <- plot_swe_anomaly(carr_mn_anom, "Carrizo")
ch_mn_anom_graph <- plot_swe_anomaly(ch_mn_anom, "Chuska")
bm_mn_anom_graph <- plot_swe_anomaly(bm_mn_anom, "Black Mesa")

# visualize next to each other

grid.arrange(carr_mn_anom_graph, ch_mn_anom_graph, bm_mn_anom_graph, heights = c(3, 3, 3))


```


### Correlations of Monthly Means Among Regions

```{r, include = FALSE}
# pearson's correlation
corr_chu_bm <- cor.test(bm_monthly$swe_mm, ch_monthly$swe_mm)
corr_chu_carr <- cor.test(carr_monthly$swe_mm, ch_monthly$swe_mm)
corr_carr_bm <- cor.test(bm_monthly$swe_mm, carr_monthly$swe_mm)

# linear models
corr_chu_bm_lm <- lm(bm_monthly$swe_mm ~ ch_monthly$swe_mm)
  chu_bm_r2 <- summary(corr_chu_bm_lm)$r.squared

corr_chu_carr_lm <- lm(carr_monthly$swe_mm ~ ch_monthly$swe_mm)
  chu_carr_r2 <- summary(corr_chu_carr_lm)$r.squared 

corr_carr_bm_lm <- lm(bm_monthly$swe_mm ~ carr_monthly$swe_mm)
  carr_bm_r2 <- summary(corr_carr_bm_lm)$r.squared 
  
# quick graph


```

Chuska vs Black Mesa R squared: `r chu_bm_r2`

Chuska vs Carrizo R squared: `r chu_carr_r2`

Black Mesa vs Carrizo R squared: `r carr_bm_r2`



## Appendix

Just to view these a bit better:

```{r, echo = FALSE}
carr_mn_anom_graph 
ch_mn_anom_graph
bm_mn_anom_graph 


```

